### 2022-08-23 design/equivalents_nips/response.tex
Commit original repos

\documentclass{article}

\usepackage{../sty/neurips2020/neurips_2020_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{verbatim}
\usepackage{lipsum}

\begin{document}

Two reviewers rated this paper very highly.
The other reviewer was anomalously low and provided a review that
lacked substance and was completely inappropriate.  
We thus ask the senior PC member to discount this review and make a
decision based on the other two thoughtful reviews, or to find
another reviewer to provide a more appropriate review.
(We will, however, address to the extent possible the comments of this reviewer below.)

\vspace{-3mm}

\paragraph{To Reviewer \#2} Thanks for the clear accept and the
helpful suggestions. In the final version, we will clarify the
discussion of $\gamma$ as a function of $k$ in Section~1.2, and we
will also expand Section~1.3 (note that a detailed discussion of
randomized iterative optimization is provided in Appendix B).

\vspace{-3mm}

\paragraph{To Reviewer \#3} Thanks for the clear accept and the helpful
suggestions. It is true that our proofs define a high probability
event, however this is merely for the analysis. The final
statements of Theorems 1 and 2 hold absolutely, rather than with high
probability. In the final version, we will expand our discussion of
low-rank approximation metrics.

\vspace{-3mm}

\paragraph{Addressing Reviewer \#4}
Below, we demonstrate that the review is inappropriate and should be
discounted, because:\vspace{-2mm}
\begin{enumerate}
\item The reviewer completely misrepresents and fails to understand
  the stated aims of the paper.
  \vspace{-1mm}
\item The reviewer makes sweeping claims which are technically wrong
  and unsupported.
\end{enumerate}\vspace{-2mm}

First of all, while we agree that the criticisms voiced by the
reviewer apply to many NeurIPS papers, this paper is \emph{not one of
  them}. We are well aware of the TCS work on low-rank
approximation and sketching. If we were to cite all of those TCS papers, it would
include hundreds of references (we choose to cite a few reviews, as is
common when an area gets to a certain level of maturity). However, more importantly,
unlike the TCS papers, in this work we are \emph{not} interested in obtaining
worst-case approximation or concentration bounds (see, e.g., lines 4
and 27). Yet, the reviewer appears to be
confused about this and states that ``the authors
show that, given a matrix A and a sketch S, the difference of the
singular values of $A^TA$  and $A^TS^TSA$ is concentrated around the
expectation.'' This is \emph{not at all} accurate in describing our
results. Instead, our goal is to provide a precise characterization,
which goes \emph{beyond worst-case bounds}, for the expected residual
projection matrix, $\mathbb{E}[I-(SA)^\dagger SA]$ (i.e., approximating an
analytically intractable deterministic quantity with a simpler
analytically tractable expression). Thus, in the context of low-rank
approximation, our goal is not
to improve on a TCS-style approximation objective (e.g., by showing a
$1+\epsilon$ error bound relative to the best rank $k$ subspace),
but rather to express the error in a simple form as a function of the
spectrum of the data matrix.  Also, unlike standard worst-case analysis, our analysis
does \emph{not} rely on satisfying some notion of the \emph{subspace
  embedding} property, which
significantly differentiates our work from that cited by the
reviewer. Note that a subspace embedding is
neither sufficient nor necessary for many numerical implementations of
sketching \citep{blendenpik,lsrn},
or statistical results
\citep{GarveshMahoney_JMLR,dobriban2019asymptotics,yang2020reduce},
as well as in the context of iterative optimization and
implicit regularization (see Section~1.3), which are discussed in
detail in the paper.

Finally, we point out the false and unsupported claims made by the
reviewer when comparing our paper to prior work.  This
likely arises from the reviewer's confusion regarding the nature of our results. (Meanwhile, the other two reviewers describe the paper as
``very well-written'' and ``clearly written''.)
Regarding \cite{optimal-matrix-product}, Reviewer \#4 states
that ``Main result here is a generalization of theorem 1'', and then
later, regarding the submission, the reviewer claims that ``the results presented here are, in
one form or another, either known results from TCS literature, or easy
corollaries''. The latter statement is incredibly broad, completely
unsupported and simply false, so we focus on the former. Regarding the former claim, Cohen et al. do provide a
low-rank approximation guarantee for sub-Gaussian sketches. However,
this result differs from ours in several respects. First of
all, instead of analyzing $\tilde A=A(SA)^\dagger SA$ directly as a
low-rank approximation with a sketch of size
$k$ (as we do), they use a larger sketch
(e.g., of size $C k/\epsilon^2$, where $C>1$ and $\epsilon< 1$) and
consider the matrix
$\tilde A _k$, defined as the best rank $k$ approximation of $\tilde
A$. This distinction is crucial for their analysis, which relies on
showing that a sketch of size sufficiently larger than $k$ ensures
a rank $k$ subspace embedding condition. This condition is not known to hold
(at least in the worst case) if the sketch is of size $k$. Our
novel analysis completely avoids subspace embeddings (which, as the reviewer points out, are central to TCS-style analysis).  This is why we can still provide
upper/lower bounds for the low-rank approximation error in this
important case. Also, the form of our bounds is completely different than that
of Cohen et al., in that we compare the error with a certain implicit
function of the singular values of $A$, which is different from the
error of the best rank $k$ approximation (used by Cohen et al.), and
so the role of $\epsilon$ in our paper is different than in
theirs. All in all, different methods are being analyzed, different types of
bounds are obtained, and completely different analysis is
used. Thus, by all indications, the reviewer is wrong, and our result is \emph{not} a corollary or a special case
of the results of Cohen et al.

\vspace{-.7\baselineskip}
\renewcommand\refname{\normalsize References}
{\scriptsize
\setlength{\bibsep}{0pt plus 0.2ex}
\bibliographystyle{abbrvnat}
\bibliography{../pap}
}

\end{document}