### 2022-08-23 design/nystrom-neurips2020-rebuttal.tex

\documentclass{article}

\usepackage{sty/neurips2020/neurips_2020_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{verbatim}
\usepackage{lipsum}
\usepackage{amsmath}
\begin{document}
We thank all the reviewers for the positive comments.

\paragraph{R1: Correctness of sr(A)-1 < k < sr(A)} 
Yes, this statement is correct. Our goal in Remark 1 is to illustrate a sharp transition in the approximation factor as $k$ approaches the stable rank from below. Remark 1(2) is an existential result i.e. there exists an $\mathbf{A}$ for which the lower bound on the approximation factor holds. 
\vspace{-2mm}
\paragraph{R1: L92 ``stable rank of A is 20" should be 40}
Here, we are referring to $\text{sr}_0(\mathbf{A})$, which is 20 and in Figure
1 it coincides with the first peak in the approximation factor.
\vspace{-2mm}
\paragraph{R1: Minor typos/comments}
Thanks for the careful read. We will make the suggested changes in the final version.
\vspace{-2mm}
\paragraph{R1: Fig 2.  Is the min $\Phi_s(k)$ line taken over $s\in\{10, 15, 25\}$ or over all values of $s\in [0,\text{rank}(\mathbf{A})]$ } The minimum is taken over all values of $s$. 
\vspace{-2mm}
\paragraph{R1: Proof of Thm 3. Importance of orthogonality of simplices}
For our setup, because of multiple simplices, the symmetry that holds for~\citet{pca-volume-sampling} breaks, and thus different subsets of the same size have different possible errors. Orthogonality, along with the choice of $\alpha_i, \beta_i$, ensures that we can pinpoint the best possible set of size $k$ and its corresponding error amongst the different choices (see proof of Lemma 4 in Appendix F). 
\vspace{-2mm}
\paragraph{R1: Other works on beyond worst case analysis}
We feel that any thorough discussion of works on ``beyond worst-case
analysis'' would have to span a very wide range of topics, which is
beyond the scope of this paper. In the final version, we will
reference \cite{roughgarden2019beyond} which provides a broader discussion on the area.
% In interest of keeping the paper precise, we did not feel that discussion of works for beyond worst case analysis on unrelated problems added anything of value to the exposition.
If the reviewer has any other works in mind which are particularly
relevant, we will definitely consider adding them.

\vspace{5mm}

\paragraph{R2: Suggested additional related works on matrix approximation}
The list of published papers on algorithms for matrix
approximation is indeed extensive. We thank the reviewer for the additional
pointers, as they are relevant to our work. In particular, we
believe that extending our results to the $\ell_p$ low-rank approximation
setting is an interesting direction for future work.

\vspace{5mm}

\paragraph{R3: Connections with overparametrization}
While we are not aware of a formal connection between the
multiple-descent phenomenon in column subset selection and
overparameterization in machine learning, both settings have to do
with obtaining a stable approximation of the data, and this stability can often be
captured by the spectral properties of the problem.
% We have established that the multiple descent phenomenon we observe is
% dependent on the spectrum of the matrix. Beyond that, there is no
% visible relationship of our work to overparameterization.

\vspace{5mm}

\paragraph{R4: Difference between k-DPP and greedy on enuite2001}
We also found the different behaviors of the approximation factor for
the two methods intriguing. One explanation for this is that the
expected approximation factor for a k-DPP is fully determined by the
eigenvalues, whereas the performance of the greedy method depends on
other characteristics of the data. However, it is not clear to us why
this is most pronounced on the eunite2001 dataset.
\vspace{-2mm}
\paragraph{R4: Correcting the effect of noisy eigenvalues}
We have not considered this, but it is indeed an
interesting question. One simple observation is that if one chooses the subset
size $k$ small enough so that the noisy tail of the spectrum consists of
only the eigenvalues with index sufficiently larger than $k$, then the effect of noise
on the approximation factor should be~minimal.

{
\setlength{\bibsep}{0pt plus 0.2ex}
\bibliographystyle{abbrvnat}
\bibliography{pap}
}
\end{document}