### 2022-08-23 design/save/colt19/rebuttal.txt

### Review 1 ###
Title: Thanks for the good suggestions

>> " 'minimax optimal experimental design'  is really 'minimax optimal among all unbiased estimators', right?"
Good point! We will clarify the section and make it clear that
we are looking for minimax optimal 'unbiased' estimators

+++MW: Not done
You only changed the title of the section.
Change title seems to be too much now?
I stressed unbiasedness in abstract

>> "...why you restrict to unbiased estimators here"
The key advantage of using unbiased estimators for the
minimax definition is that we can say that the least
squares estimator w_LS is the minimum variance estimator
among them (Proposition 7). If we include all biased
estimators, then it is not clear what w_LS
should be replaced with to even define the minimax value. Also, unbiased estimators 
are closed under averaging and this immediately lets us
reduce the variance/loss. We will add more discussion to clarify this.

+++MW: not done
Added a minimal remark before Prop. 7 in Magenta

>> "how does this compare to the pseudo-sequential setting"
This setting gives more freedom to the design, so we would expect that the bounds should be at least as good. Certainly, this is a great direction for future work!

>> "I found it confusing that \hat{\bf w}({\bf y}_S ) is not written conditional on {\bf X}"
Good point! Of course the estimator function has to depend on X. Since X is fixed beforehand, and then w_hat is chosen afterwards, we treat the dependence on X as implicit in the choice of w_hat. We will clarify this point in the final version!

+++MW: not done
Sometimes with condition on X and sometimes we dont
Indeed confusing. See for example statement of Thm 4
Would be hard to fix though

hat w (y_S)   but  w_LS(y|X)

### Review 2 ###
Title: The minimax part is not a distraction

Thanks for all the comments and suggestions!
Yes - the paper is a bit overbearing and we understand that
you want us to simplify it. However Prop. 9 is not a minor
result. We will highlight this proposition more clearly.

>> "The authors define the minimax-optimal value in Definition 8 and provide some minor results in Prop 9."
>> "Such a ``minimax'' discussion acts as a distraction"
>> "make the paper less daunting (in terms of the number of results/lemmas)."
Reasons for keeping Prop. 9:
Almost all of the claims made in the paper are needed to obtain Prop. 9
and it provides BOTH upper and lower bounds for the minimax value Rk*. 
The bounds are tight up to within a constant factor. 
We will make this point clearer and the whole section more approachable in the final version.

+++MW: done

### Review 3 ###
Title: We have upper and lower bounds for the minimax value

>> "My major concern is with Sec. 1.2, and the so-called “minimax optimality” "
>> "For a typical paper, the next step would then be establishing matching upper and lower bounds of Rk*(X),"
BOTH upper and lower bounds for Rk* are given in Prop. 9 (part 2 and 3). 
The bounds are tight up to within a constant factor. 
In fact, the upper bound of C \phi/k (part 2) is a consequence of our main result, Theorem 4. 
Thus, while our main proposed estimator does not obtain exact minimax optimality 
(this appears very hard to attain), it does lead to a very strong guarantee for Rk*.

>> "The guarantees ... are additive, which are very different from the approximation
>> guarantees from Allen-Zhu et al. (2017), Wang et al. (2017) and Nikolov et al
>> (2018), deriving multiplicative guarantees."
>> "I feel this difference should be highlighted and discussed."
We completely agree. While we already point this out in the related work, we will highlight it further in the final version.
+++MW: not done
Dont know how to do this

>> "I suggest avoiding the usage of the term “minimax”, while replacing it with “minimum variance unbiased” "
>> "I suggest deriving some example scalings of \phi assuming X is randomly generated from some underlying distribution"
Good suggestions, thank you! We will address them in the final version.