### 2022-08-23 design/continuous_colt/colt2020-sample.tex

\documentclass[anon]{colt2020} % Anonymized submission
% \documentclass{colt2020} % Include author names

% The following packages will be automatically loaded:
% amsmath, amssymb, natbib, graphicx, url, algorithm2e

\title[Short Title]{Full Title of Article}
\usepackage{times}
% Use \Name{Author Name} to specify the name.
% If the surname contains spaces, enclose the surname
% in braces, e.g. \Name{John {Smith Jones}} similarly
% if the name has a "von" part, e.g \Name{Jane {de Winter}}.
% If the first letter in the forenames is a diacritic
% enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}

% Two authors with the same address
% \coltauthor{\Name{Author Name1} \Email{abc@sample.com}\and
%  \Name{Author Name2} \Email{xyz@sample.com}\\
%  \addr Address}

% Three or more authors with the same address:
% \coltauthor{\Name{Author Name1} \Email{an1@sample.com}\\
%  \Name{Author Name2} \Email{an2@sample.com}\\
%  \Name{Author Name3} \Email{an3@sample.com}\\
%  \addr Address}

% Authors with different addresses:
\coltauthor{%
 \Name{Author Name1} \Email{abc@sample.com}\\
 \addr Address 1
 \AND
 \Name{Author Name2} \Email{xyz@sample.com}\\
 \addr Address 2%
}

\begin{document}

\maketitle

\begin{abstract}%
  This is a great paper and it has a concise abstract.%
\end{abstract}

\begin{keywords}%
  List of keywords%
\end{keywords}

\section{Introduction}

This is where the content of your paper goes.
\begin{itemize}
  \item Limit the main text (not counting references and appendices) to 12 PMLR-formatted pages, using this template.
  \item Include, either in the main text or the appendices, all details, proofs and derivations required to substantiate the results.
  \item Include {\em in the main text} enough details, including proof details, to convince the reviewers of the contribution, novelty and significance of the submissions.
  \item Use the \textbackslash documentclass[anon]{colt2020} option during submission process -- 
this automatically hides the author names listed under \textbackslash coltauthor. 
Do not include author names in the remainder of the text, and to the extent possible, avoid directly identifying the authors. 
You should still include all relevant references, including your own, and any other relevant discussion, 
even if this might allow a reviewer to infer the author identities. 
Use \textbackslash documentclass[final]{colt2020} only during camera-ready submission.
\end{itemize}

% Acknowledgments---Will not appear in anonymized version
\acks{We thank a bunch of people.}

\bibliography{yourbibfile}

\appendix

\section{My Proof of Theorem 1}

This is a boring technical proof.

\section{My Proof of Theorem 2}

This is a complete version of a proof sketched in the main text.

\end{document}