### 2022-08-23 design/continuous_nips/continuous.tex

 \newif\ifisarxiv
 \isarxivfalse

 \ifisarxiv
 \documentclass[11pt]{article}
 \usepackage[round]{natbib}
 \usepackage{fullpage}
 \usepackage{amsmath}
 \usepackage{amssymb}
 \usepackage{amsthm}
 \usepackage{mathabx}
 \usepackage{color}
 \usepackage{cancel}
 \usepackage{wrapfig}
 \usepackage{subfigure}
 \usepackage{caption}
 \usepackage{xfrac}
 \usepackage{algorithm}
 \usepackage{algorithmic}
 \usepackage{mdframed}
 \usepackage{hyperref}
 \usepackage{xcolor}
 \hypersetup{
 	colorlinks,
 	linkcolor={red!40!gray},
 	citecolor={blue!40!gray},
 	urlcolor={blue!70!gray}
 }
 \usepackage{cleveref}

 \title{Exact expressions for double descent
 and implicit regularization
 \\
 via surrogate random design}
   \author{%
           \textbf{Micha{\l } Derezi\'{n}ski} \\
   Department of Statistics\\
   University of California, Berkeley\\
   \texttt{mderezin@berkeley.edu}\\
   \and
   \textbf{Feynman Liang} \\
   Department of Statistics\\
   University of California, Berkeley\\
   \texttt{feynman@berkeley.edu}
   \and
    \textbf{Michael W. Mahoney}\\
   ICSI and Department of Statistics\\
   University of California, Berkeley\\
   \texttt{mmahoney@stat.berkeley.edu}
 }

 %  \newcommand{\jmlrBlackBox}{\rule{1.5ex}{1.5ex}}
   \providecommand{\BlackBox}{\rule{1.5ex}{1.5ex}}
 %  \newcommand{\jmlrQED}{\hfill\jmlrBlackBox\par\bigskip}
   \renewenvironment{proof}%
   {%
    \par\noindent{\bfseries\upshape Proof\ }%
   }%
 {\hfill\BlackBox\par\bigskip}

\else

\documentclass{article}
\usepackage[final]{neurips_2020}

\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsmath}
\usepackage{cancel}
\usepackage{cleveref}
\usepackage{subfigure}
\usepackage{todonotes}
 \usepackage{wrapfig}
\usepackage{xcolor}
\usepackage{xfrac}
\hypersetup{
	colorlinks,
	linkcolor={red!40!gray},
	citecolor={blue!40!gray},
	urlcolor={blue!70!gray}
}

\title{Exact expressions for double descent and implicit regularization via
surrogate random design}
  \author{%
Micha{\l } Derezi\'{n}ski \\
  Department of Statistics\\
  University of California, Berkeley\\
  \texttt{mderezin@berkeley.edu}\\
  \And
Feynman Liang \\
  Department of Statistics\\
  University of California, Berkeley\\
  \texttt{feynman@berkeley.edu}
  \And
Michael W. Mahoney\\
  ICSI and Department of Statistics\\
  University of California, Berkeley\\
  \texttt{mmahoney@stat.berkeley.edu}
}
% \usepackage{times}

% \coltauthor{%
%   \Name{Micha{\l } Derezi\'{n}ski} \Email{mderezin@berkeley.edu}\\
%   \addr Department of Statistics\\
%   University of California, Berkeley\\
%   \AND
%   \Name{Feynman Liang} \Email{feynman@berkeley.edu}\\
%   \addr Department of Statistics\\
%   University of California, Berkeley\\
%   \AND
%   \Name{Zhenyu Liao} \Email{TODO}\\
%   \addr Department of Statistics\\
%   University of California, Berkeley\\
%   \AND
%   \Name{Michael W. Mahoney} \Email{mmahoney@stat.berkeley.edu}\\
%   \addr ICSI and Department of Statistics\\
%   University of California, Berkeley
% }

% \usepackage{float}
% \usepackage{wrapfig}

\fi

\input{../shortdefs}

\newcommand{\RED}{\color[rgb]{0.70,0,0}}

\begin{document}
\maketitle

\begin{abstract}
  Double descent refers to the phase transition that is exhibited by
  the generalization error of unregularized learning models when varying the ratio
  between the number of parameters and the number of training
  samples. The recent success of highly over-parameterized machine learning
  models such as deep neural networks has motivated a theoretical analysis of
  the double descent phenomenon in classical models such as linear
  regression which can also generalize well in the over-parameterized
  regime. We provide the first exact non-asymptotic
  expressions for double descent of the minimum norm linear
  estimator. Our approach involves constructing a special
  determinantal point process  which we call surrogate random
  design, to replace the standard i.i.d.~design of the training
  sample. This surrogate design admits exact expressions for the mean
  squared error of the estimator while preserving the key properties
  of the standard design. We also establish an exact implicit
  regularization result for over-parameterized training samples. In
  particular, we show that, for the surrogate design, the implicit bias
  of the unregularized minimum norm estimator precisely corresponds to
  solving a ridge-regularized least squares problem on the population
  distribution. In our analysis we introduce a new mathematical tool of
  independent interest: the class of random matrices for which
  determinant commutes with expectation.
\end{abstract}

\section{Introduction}

Classical statistical learning theory asserts that to achieve generalization
one must use training sample size that sufficiently exceeds the complexity of
the learning model, where the latter is typically represented by the number of
parameters \citep[or some related structural parameter; see][]{HFT09}.  In particular,
this seems to suggest the conventional wisdom that one should not use models
that fit the training data exactly.  However, modern machine learning practice
often seems to go against this intuition, using models with so many parameters
that the training data can be perfectly interpolated, in which case the
training error vanishes. It has been shown that models such as deep neural
networks, as well as certain so-called interpolating kernels and decision
trees, can generalize well in this regime. In particular,
\cite{BHMM19} empirically demonstrated a phase transition in generalization
performance of learning models which occurs at an \emph{interpolation
  thershold}, i.e., a point where training error goes to zero (as one varies the
ratio between the model complexity and the sample size). Moving away from this
threshold in either direction tends to reduce the generalization error, leading
to the so-called \emph{double descent} curve.

To understand this
surprising phenomenon, in perhaps the simplest possible setting, we
study it in the context of linear or least squares regression.
Consider a full rank $n\times d$ data matrix $\X$ and a vector $\y$ of
responses corresponding to each of the $n$ data points (the rows of $\X$), where we wish to
find the best linear model $\X\w\approx \y$, parameterized by a
$d$-dimensional vector $\w$.
The simplest example of an estimator that has been shown to exhibit
the double descent phenomenon \citep{belkin2019two} is the
Moore-Penrose estimator, $\wbh=\X^\dagger\y$:
in the so-called over-determined regime, i.e., when $n>d$, it corresponds to the
least squares solution, i.e., $\argmin_{\w} \|\X\w-\y\|^2$; and in the
under-determined regime (also known as
over-parameterized or interpolating), i.e., when $n<d$, it
corresponds to the minimum norm solution to the linear system $\X\w=\y$.
Given the ubiquity of linear regression and the Moore-Penrose
solution, e.g., in kernel-based machine learning, studying the
performance of this estimator can shed some light on the effects of
over-parameterization/interpolation in machine learning more generally.
Of particular interest are results that are exact (i.e., not upper/lower bounds) and
non-asymptotic (i.e., for large but still finite $n$ and~$d$).

We build on methods from Randomized Numerical Linear Algebra (RandNLA) in order
to obtain \emph{exact non-asymptotic expressions} for the mean squared error
(MSE) of the Moore-Penrose estimator (see Theorem~\ref{t:mse}).  This provides
a precise characterization of the double descent phenomenon for the
linear regression problem. In obtaining these results,
we are able to provide precise formulas for the \emph{implicit regularization}
induced by minimum norm solutions of under-determined training samples,
relating it to classical ridge regularization (see Theorem~\ref{t:unbiased}).
% This result has been observed empirically for RandNLA
% methods~\citep{Mah-mat-rev_JRNL}, but it has also been shown in deep
% learning~\citep{Ney17_TR} and machine learning~\citep{Mah12} more generally.
To obtain our precise results, we use a somewhat non-standard random
design, based on a specially chosen determinantal point process (DPP), which we
term surrogate random design. DPPs are a family of non-i.i.d.~sampling
distributions which are typically used to induce diversity in the
produced samples \citep{dpp-ml}. Our aim in using a DPP as a surrogate
design is very different: namely, to make certain quantities (such as the MSE)
analytically tractable, while accurately \emph{preserving} the underlying
properties of the original data distribution. This strategy might seem
counter-intuitive since DPPs are typically found most useful when they
\emph{differ} from the data distribution. However, we show both theoretically
(Theorem~\ref{t:asymptotic}) and empirically
(Section~\ref{sec:asymp-conj-details}), that for many commonly studied data
distributions, such as multivariate Gaussians, our DPP-based surrogate
design accurately preserves the key properties of the standard i.i.d.~design
(such as the MSE), and even matches it exactly in the high-dimensional
asymptotic limit. In our analysis of the surrogate design, we
introduce the concept of \emph{determinant preserving
  random matrices} (Section \ref{s:dp}), a class of random matrices for which determinant
commutes with expectation, which should be of independent interest.

%  To
% obtain our precise results, we use a somewhat non-standard random design, which
% we term surrogate random design (see Section~\ref{s:determinantal} for a
% detailed discussion), and which we expect will be of more general interest.
% Informally, the goal of a surrogate random design is to modify an original
% design to capture its main properties while being ``nicer'' in some useful way.
% In Theorem~\ref{t:asymptotic} and Appendix \ref{sec:asymp-conj-details} we show, both
% theoretically and empirically, that our surrogate design accurately preserves
% the key properties of the original design when the data distribution is a
% multivariate Gaussian.

\subsection{Main results: double descent and implicit regularization}

As the performance metric in our analysis, we use the \emph{mean
  squared error} (MSE), defined as
$\mathrm{MSE}[\wbh]=\E\big[\|\wbh-\w^*\|^2\big]$, where $\w^*$ is a fixed
underlying linear model of the responses.
In analyzing the MSE, we make the following standard assumption that
the response noise is homoscedastic.
\begin{assumption}[Homoscedastic noise]\label{a:linear}
  The noise $\xi =y(\x)- \x^\top\w^*$ has mean $0$ and variance $\sigma^2$.
\end{assumption}

\noindent
Our main result provides an exact expression for the MSE of the
Moore-Penrose estimator under our surrogate design denoted $\Xb\sim S_\mu^n$, where
$\mu$ is the $d$-variate distribution of the row vector $\x^\top$ and $n$ is the sample
size. This surrogate is
used in place of the standard $n\times d$ random design $\X\sim\mu^n$, where $n$
data points (the rows of $\X$) are sampled independently from
$\mu$. We form the surrogate by constructing a determinantal point
process with $\mu$ as the background measure, so that $S_\mu^n(\X)\propto
  \pdet(\X\X^\top)\mu(\X)$, where $\pdet(\cdot)$ denotes the pseudo-determinant
(details in Section~\ref{s:determinantal}).  Unlike for the standard
design, our MSE formula is fully expressible as a function of the covariance
matrix $\Sigmab_\mu=\E_\mu[\x\x^\top]$. To state our main result, we
need an additional minor assumption on $\mu$ which is satisfied by
most standard continuous distributions (e.g., multivariate Gaussians).
\begin{assumption}[General position]\label{a:general-position}
  For $1\leq n \leq d$, if $\X\sim\mu^n$, then $\rank(\X)=n$ almost surely.
\end{assumption}

\noindent
Under Assumptions~\ref{a:linear} and~\ref{a:general-position}, we can
establish our first main result, stated as the following theorem, where
we use $\X^\dagger$ to denote the Moore-Penrose inverse of $\X$.
% , which gives an exact non-asymptotic
% expression for the MSE of the Moore-Penrose estimator  the
% surrogate design $S_\mu^n$.

\begin{theorem}[Exact non-asymptotic MSE]
  \label{t:mse}
  If the response noise is homoscedastic % with variance~$\sigma^2$
  (Assumption~\ref{a:linear}) and $\mu$ is in
  general position (Assumption~\ref{a:general-position}), then for
  any $\w \in \mathbb{R}^d$,
  $\Xb\sim S_\mu^n$ (Definition \ref{d:surrogate}), and
  $\yb_i=y(\xbb_i)$, %we have
  \vspace{-2mm}
  \begin{align*}
    \MSE{\Xb^\dagger\ybb} =
    \begin{cases}
      \sigma^2\,\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)\cdot
      \frac{1-\alpha_n}{d-n}\ +\
      \frac{\w^{*\top}(\Sigmab_\mu+\lambda_n\I)^{-1}\w^*}
      {\tr((\Sigmab_\mu+\lambda_n\I)^{-1})}\cdot (d-n),
                                                                 & \text{for }n<d, \\
      \sigma^2\, \tr(\Sigmab_\mu^{-1}),                          & \text{for }n=d, \\
      \sigma^2\,\tr(\Sigmab_\mu^{-1})\cdot\frac{1-\beta_n}{n-d}, & \text{for
      }n>d,
    \end{cases}
  \end{align*}
  with $\lambda_n\geq 0$ defined by
  $n=\tr(\Sigmab_\mu (\Sigmab_\mu+\lambda_n\I)^{-1})$, \
  $\alpha_n=\det(\Sigmab_\mu(\Sigmab_\mu+\lambda_n\I)^{-1})$
  and $\beta_n=\ee^{d-n}$.
\end{theorem}
\begin{definition}\label{d:expression}
  We will use $\Mc = \Mc(\Sigmab_\mu, \w^*,\sigma^2,n)$ to denote the above expressions
  for $\MSE{\Xb^\dagger\ybb}$.
\end{definition}

\begin{figure}[t]
  \centering
  \subfigure[%
    Surrogate MSE expressions (Theorem \ref{t:mse}) closely match
    numerical estimates even for non-isotropic
    features. Eigenvalue decay leads to a steeper
    descent curve in the under-determined regime ($n<d$).]{%
    \includegraphics[width=0.48\textwidth]{figs/descent-intro}
  }
  \hfill
  \subfigure[%
    The mean of the estimator $\X^\dagger\y$ exhibits
    shrinkage which closely matches the shrinkage of a
    ridge-regularized least squares optimum (theory lines), as characterized by
    Theorem \ref{t:unbiased}.]{%
    \includegraphics[width=0.48\textwidth]{figs/descent-shrinkage}
  }
  \caption{Illustration of the main results for $d=100$ and
    $\mu=\Nc(\zero,\Sigmab)$ where $\Sigmab$ is diagonal with
    eigenvalues decaying exponentially and scaled so that
    $\tr(\Sigmab^{-1})=d$. We use our surrogate
    formulas to plot (a) the MSE (Theorem \ref{t:mse}) and (b) the norm of the expectation (Theorem
    \ref{t:unbiased}) of the Moore-Penrose estimator (\emph{theory}
    lines), accompanied by the empirical estimates based on the standard
    i.i.d.~design (error bars are three times the standard error of the
    mean). We consider three different condition numbers $\kappa$ of
    $\Sigmab$, with \emph{isotropic} corresponding to $\kappa=1$,
    i.e., $\Sigmab=\I$. We use $\sigma^2=1$ and
    $\w^*=\frac1{\sqrt{d}}\one$.}
  %\vspace{-5mm}
  \label{f:intro}
\end{figure}

\noindent
Proof of Theorem \ref{t:mse} is given in Appendix \ref{a:mse-proof}.
For illustration, we plot the MSE expressions in Figure~\ref{f:intro}a,
comparing them with empirical estimates of the true MSE under the
i.i.d.~design for a multivariate Gaussian distribution
$\mu=\Nc(\zero,\Sigmab)$ with several different covariance matrices $\Sigmab$. We keep the number of features $d$ fixed to
$100$ and vary the number of samples $n$, observing a double descent
peak at $n=d$. We observe that our theory aligns well with
the empirical estimates, whereas
previously, no such theory was available except for special
cases such as $\Sigmab=\I$ (more details in Theorem~\ref{t:asymptotic}
and Section~\ref{sec:asymp-conj-details}). The plots
show that varying the spectral decay of $\Sigmab$ has a significant effect on the
shape of the curve in the under-determined regime. We use the
horizontal line to denote the MSE of the null estimator
$\mathrm{MSE}[\zero]=\|\w^*\|^2=1$. When the eigenvalues of $\Sigmab$
decay rapidly, then the Moore-Penrose estimator suffers less error
than the null estimator for some values of $n<d$, and the curve
exhibits a local optimum in this regime.

One important aspect of Theorem~\ref{t:mse} comes from the relationship between $n$ and the parameter $\lambda_n$, which together satisfy $n=\tr(\Sigmab_\mu (\Sigmab_\mu+\lambda_n\I)^{-1})$.
This expression is precisely the classical notion of \emph{effective
  dimension} for ridge regression regularized with
$\lambda_n$~\citep{ridge-leverage-scores}, and it arises here even though there is
no explicit ridge regularization in the problem being considered in
Theorem~\ref{t:mse}.
The global solution to the ridge regression task (i.e., $\ell_2$-regularized
least squares) with parameter $\lambda$ is defined as:
\begin{align*}
  \argmin_\w \Big\{\E_{\mu,y}\big[\big(\x^\top\w-y(\x)\big)^2\big]
  + \lambda\|\w\|^2\Big\}\ =\ (\Sigmab_\mu +
  \lambda\I)^{-1}\v_{\mu,y},\quad\text{where } \v_{\mu,y}=\E_{\mu,y}[y(\x)\,\x ].
\end{align*}
When Assumption \ref{a:linear} holds, then
$\v_{\mu,y}=\Sigmab_\mu\w^*$, however ridge-regularized least squares
is well-defined for much more general response models.
Our second result makes a direct connection between the (expectation
of the) unregularized minimum norm solution on the sample
and the global ridge-regularized solution.
While the under-determined regime (i.e., $n<d$) is of primary interest to us,
for completeness we state this result for arbitrary values of $n$ and $d$.
Note that, just like the definition of regularized least squares, this
theorem applies more generally than Theorem~\ref{t:mse}, in that it
does \emph{not} require the responses to follow any linear model as in
Assumption~\ref{a:linear} (proof in Appendix~\ref{s:unbiased-proof}).
\begin{theorem}[Implicit regularization of Moore-Penrose estimator]
  \label{t:unbiased}
  For $\mu$ satisfying
  % \footnote{The proof of Theorem \ref{t:unbiased}
  %   can be easily extended to probability measures $\mu$ that do not
  %   satisfy Assumption~\ref{a:general-position} (such as discrete distributions). We include this
  %   assumption here to simplify the presentation.}
  Assumption~\ref{a:general-position} and
  $y(\cdot)$ s.t.~$\v_{\mu,y}=\E_{\mu,y}[y(\x)\,\x]$ is well-defined,
  $\Xb\sim S_\mu^n$ (Definition \ref{d:surrogate}) and $\yb_i=y(\xbb_i)$,
  \begin{align*}
    \E\big[\Xb^\dagger\ybb\big] =
    \begin{cases}
      (\Sigmab_\mu + \lambda_n\I)^{-1}\v_{\mu,y} & \text{for }n<d,     \\
      \Sigmab_\mu^{-1}\v_{\mu,y}                 & \text{for }n \ge d,
    \end{cases}
  \end{align*}
  where, as in Theorem \ref{t:mse}, $\lambda_n$ is such that the effective dimension
  $\tr(\Sigmab_\mu(\Sigmab_\mu+\lambda_n\I)^{-1})$ equals $n$.
\end{theorem}

\noindent
That is, when $n < d$, the Moore-Penrose estimator (which itself is
not regularized), computed on the
random training sample, in expectation equals the global ridge-regularized least
squares solution of the underlying regression
problem. Moreover, $\lambda_n$, i.e., the amount
of implicit $\ell_2$-regularization, is controlled by the degree of
over-parameterization in such a way as to ensure that $n$ becomes the ridge effective dimension
(a.k.a.~the effective degrees of freedom).

We illustrate this result in Figure
\ref{f:intro}b, plotting the norm of the expectation of the
Moore-Penrose estimator. As for the MSE, our surrogate theory aligns
well with the empirical estimates for i.i.d.~Gaussian designs, showing
that the shrinkage of the unregularized estimator in the
under-determined regime matches the implicit
ridge-regularization characterized by Theorem \ref{t:unbiased}. While the shrinkage
is a linear function of the sample size $n$ for isotropic features
(i.e., $\Sigmab=\I$), it
exhibits a non-linear behavior for other spectral decays.
Such \emph{implicit regularization} has been studied
previously~\citep[see, e.g.,][]{MO11-implementing, %PM11, GM14_ICML,
  Mah12}; it has
been observed empirically for RandNLA sampling
algorithms~\citep{MMY15}; and it has also received attention more
generally within the context of neural networks~\citep{Ney17_TR}. While our implicit regularization result
is limited to the Moore-Penrose estimator, this new connection (and
others, described below) between the minimum norm solution of an unregularized
under-determined system and a ridge-regularized least squares solution
offers a simple interpretation for the implicit regularization
observed in modern machine learning architectures.

Our exact non-asymptotic expressions in Theorem~\ref{t:mse} and
our exact implicit regularization results in Theorem~\ref{t:unbiased}
are derived for the surrogate design, which is a
non-i.i.d.~distribution based on a determinantal point process. However, Figure \ref{f:intro}
suggests that those expressions accurately describe the MSE (up to lower order
terms) also under the standard i.i.d.~design $\X\sim\mu^n$ when $\mu$ is a multivariate Gaussian.
As a third result, we verify that the surrogate expressions for the
MSE are asymptotically consistent with the MSE of an i.i.d.~design,
for a wide class of distributions which include multivariate Gaussians.

\begin{theorem}[Asymptotic consistency of surrogate design]
  \label{t:asymptotic}
  Let $\X \in \mathbb{R}^{n \times d}$ have i.i.d.~rows $\x_i^\top = \z_i^\top\Sigmab^{\frac12}$
  where $\z_i$ has independent zero mean and unit variance
  sub-Gaussian entries, and suppose that Assumptions~\ref{a:linear}
  and \ref{a:general-position} are satisfied. Furthermore, suppose
  that there
  exist $c, C, C^* \in \mathbb{R}_{> 0}$ such that
  $C \I \succeq \Sigmab \succeq c \I \succ 0$ and $\| \w^* \| \le
    C^*$. Then
  %and the limiting spectral measure (i.e., normalized counting measure of the eigenvalues) of $\Sigmab \in \mathbb R^{d \times d}$ converges weakly to some measure $\nu$ as $d \to \infty$. Then
  \begin{align*}
    \MSE{\X^\dagger\y} - \Mc(\Sigmab, \w^*,\sigma^2,n) \to 0
  \end{align*}
  with probability one as $d,n \to \infty$ with $n/d \to \bar c \in (0,\infty) \setminus \{1\}$.
  % Let $\rho=n/d\neq 1$, $\X\sim \mu^n$
  % and $y_i=y(\x_i)$ satisfy Assumption~\ref{a:linear}. If $d\geq
  % c_\rho=\frac3{|1-\rho|}$ and
  % \begin{align*}
  %   \mu &= \begin{cases}
  %     \Nc(\zero,\I)&\text{when }n<d-1,\\
  %     \Nc(\zero,\Sigmab),\ \Sigmab\succ\zero &\text{when }n>d+1,
  %   \end{cases}
  % \end{align*}
  % then the absolute difference between surrogate
  % expressions and the true MSE is bounded as follows:
  % % &\text{if}\qquad  (1)\quad \rho<1\ \text{ and }\ \mu=\Nc(\zero,\I)\qquad\text{or}\qquad
  % %     (2)\quad \rho > 1\ \text{ and }\ \mu=\Nc(\zero,\Sigmab),\
  % %     \Sigmab\succ\zero,\\
  % \begin{align*}
  % \Big|\MSE{\X^\dagger\y}-\Mc(\Sigmab, \w^*,\sigma^2,n)\Big|\ \leq\
  % \frac{c_\rho}{d} \cdot \Mc(\Sigmab,\w^*,\sigma^2,n).
  % \end{align*}
\end{theorem}

% \begin{remark}
% For $n$ equal to $d-1$, $d$ or $d+1$, the true MSE under Gaussian
% random design can be infinite, whereas the
% surrogate MSE is finite and has a closed form expression.
% \end{remark}

\noindent
The above result is particularly remarkable since our surrogate design
is a determinantal point process. DPPs are commonly used in ML to
ensure that the data points in a sample are well spread-out. However,
if the data distribution is sufficiently regular (e.g., a multivariate
Gaussian), then the i.i.d.~samples are already spread-out reasonably
well, so rescaling the distribution by a determinant has a negligible
effect that vanishes in the high-dimensional regime.
Furthermore, our empirical estimates (Figure~\ref{f:intro}) suggest that the surrogate expressions
are accurate not only in the asymptotic limit, but even for moderately large
dimensions. Based on a detailed empirical analysis described in
Section \ref{sec:asymp-conj-details},
we conjecture that the convergence described in
Theorem~\ref{t:asymptotic} has the rate of~$O(1/d)$.
% holds true in the under-determined regime without the assumption that
% $\Sigmab=\I$ (see Conjectures \ref{c:wishart} and
% \ref{c:projection}). In this case, no formula is known for 
% $\MSE{\X^\dagger\y}$, whereas the expressions for the surrogate
% Gaussian design naturally extend.

% \subsection{Key techniques: surrogate designs and
%   determinant preserving matrices}

% The standard random design model for linear regression assumes that
% each pair $(\x_i^\top,y_i)$ is drawn independently, where the row vector
% $\x_i^\top$ comes from some $d$-variate distribution $\mu$ and $y_i=y(\x_i)$ is a random response variable drawn conditionaly on $\x_i$.
% Precise theoretical analysis of under-determined regression in this
% setting poses significant challenges, even in such special cases as
% the Moore-Penrose estimator and a Gaussian data distribution $\mu$.
% Rather than trying to directly analyze the usual i.i.d.~random design
% $\X\sim\mu^n$ described above, we modify it slightly by introducing
% the notion of a \emph{surrogate random design}, $\Xb\sim S_\mu^n$.
% Informally, the goal of a surrogate random design is to modify an
% original design to capture the main properties of the original design,
% while being ``nicer'' for theoretical or empirical analysis.
% In particular, here, we will modify the distribution of matrix $\X$ so
% as to:
% \begin{enumerate}
%   \item closely preserve the behavior of the Moore-Penrose estimator
% from the i.i.d.~design; and
%   \item obtain exact expressions for double descent in terms of the
%     mean squared error.
% \end{enumerate}
% A key element in the construction of our surrogate designs involves rescaling the measure $\X\sim\mu^n$ by the pseudo-determinant $\pdet(\X\X^\top)$, i.e., a product of the non-zero eigenvalues.
% A similar type of determinantal design was suggested by
% \cite{correcting-bias-journal}, but it was restricted there only to
% $n\geq d$.
% We broaden this definition by not only allowing the sample size to be
% less than $d$, but also allowing it to be randomized.
% A key property of our determinantal design matrix $\Xb$ is that
% the expectation of a function $F(\Xb)$ can be, up to normalization,
% expressed as:
% \begin{align*}
%   \E[F(\Xb)]\ \propto\ \E[\pdet(\X\X^\top)F(\X)],
%   \end{align*}
% where $\X\sim\mu^K$ and $K$ is a random variable (see details in Definition~\ref{d:det}).
% Then, we define (in Definition~\ref{d:surrogate}) our surrogate design $S_\mu^n$ for each $n>0$ as a determinantal design with a carefully chosen random variable $K$, so that the expected sample size is equal to $n$ and so that it is possible to derive closed form expressions for the MSE.
% We achieve this by using the Poisson distribution to construct the variable~$K$.

% The key technical contribution that allows us to derive the MSE for determinantal designs is the concept of \emph{determinant preserving random matrices}, a notion that we expect to be useful more generally.
% Specifically, in Section~\ref{s:dp} we define a class of $d\times d$ random
% matrices $\A$ for which taking the determinant commutes with taking
% the expectation, for the matrix itself and any of its square submatrices (see Definition~\ref{d:main}):
% \begin{align*}
%   \E\big[\!\det(\A_{\Ic,\Jc})\big] =
%   \det\!\big(\E[\A_{\Ic,\Jc}]\big)\quad \text{for all }\Ic,\Jc\subseteq
%   [d]\text{ s.t. }|\Ic|=|\Jc|.
% \end{align*}
%   Not all random matrices satisfy this property, however many
% interesting and non-trivial examples can be found. Constructing
% these examples is facilitated by the closure properties that this
% class enjoys. In particular, if $\A$ and $\B$ are determinant
% preserving and independent, then $\A+\B$ and $\A\B$ are also
% determinant preserving (see Lemma \ref{t:ring}). We use these
% techniques to prove a number of determinantal expectation
% formulas needed in obtaining Theorems~\ref{t:mse}
% and~\ref{t:unbiased}. For example, we show that if $\X\sim\mu^K$,
% where $K$ is a Poisson random variable, then:
% \begin{align*}
%   \text{(Lemma \ref{l:poisson})}\quad  \E\big[\det(\X^\top\X)\big]
%   &= \det\!\big(\E[\X^\top\X]\big),\\
% \text{(Lemma \ref{l:normalization})}\quad  \E\big[\det(\X\X^\top)\big]
% &= \ee^{-\E[K]}\det\!\big(\I + \E[\X^\top\X]\big).
% \end{align*}

\section{Related work}
\label{s:related-work}

There is a large body of related work, which for simplicity we cluster into three groups.

\textbf{Double descent.}
The double descent phenomenon has been observed empirically
in a number of learning models, including neural networks
\citep{BHMM19,GJSx19_TR}, kernel methods \citep{BMM18_TR,BRT18_TR},
nearest neighbor models \citep{BHM18_TR}, and decision trees \citep{BHMM19}. The
theoretical analysis of double descent, and more broadly the generalization
properties of interpolating estimators, have primarily focused on various forms of
linear regression \citep{BLLT19_TR,LR18_TR,HMRT19_TR,MVSS19_TR}. Note
that while we analyze the classical mean squared error, many
works focus on the squared prediction error.
Also, unlike in our work, some of the literature on double descent deals with linear regression in the
so-called \emph{misspecified} setting, where the set of observed
features does not match the feature space in
which the response model is linear
\citep{belkin2019two,HMRT19_TR,Mit19_TR,MM19_TR}, e.g., when the
learner observes a random subset of $d$ features from a larger
population.

\begin{wrapfigure}{r}{\ifisarxiv 0.5\else 0.44\fi\textwidth}
  \ifisarxiv\else \vspace{-.2cm}\fi
  \centering
  \includegraphics[width=.43\textwidth]{figs/descent-model-log}
  \ifisarxiv\else \vspace{-3mm}\fi
  \caption{
    Surrogate MSE as a function of $d/n$, with $n$
    fixed to $100$ and varying dimension $d$ and condition number
    $\kappa$, for signal-to-noise ratio $\mathrm{SNR}=1$.}
  \label{f:model}
  \ifisarxiv\else  \vspace{-.5cm}\fi
\end{wrapfigure}

The most directly comparable to our setting is the recent work of
\cite{HMRT19_TR}. They study how
varying the feature dimension affects the (asymptotic) generalization
error for linear regression, % for both well-specified and misspecified tasks
however their analysis is limited to certain special settings such as an
isotropic data distribution. As an additional
point of comparison, in Figure~\ref{f:model} we plot the MSE
expressions of Theorem~\ref{t:mse} when varying the
feature dimension~$d$ (the setup is the same as in
Figure~\ref{f:intro}). Our plots follow the
trends outlined by \cite{HMRT19_TR} for the isotropic case (see their Figure~2),
but the spectral decay of the covariance (captured by our new
MSE expressions) has a significant effect on the descent curve.
This leads to generalization in the under-determined regime even when
the signal-to-noise ratio ($\mathrm{SNR}=\|\w^*\|^2/\sigma^2$) is 1,
unlike suggested by \cite{HMRT19_TR}.

\textbf{RandNLA and DPPs.}
Randomized Numerical Linear Algebra
\citep{DM16_CACM,RandNLA_PCMIchapter_TR} has
traditionally focused on obtaining purely algorithmic improvements for tasks
such as least squares regression, %  and low-rank approximation via techniques that include
% sketching \citep{sarlos-sketching} and i.i.d.~leverage score sampling
% \citep{Drineas2006sampling}. However,
but there has been growing
interest in understanding the statistical properties of these
randomized methods \citep{MMY15,GarveshMahoney_JMLR} and
a beyond worst-case analysis \cite{precise-expressions}.
Determinantal point processes \citep{dpps-in-randnla,dpp-ml} have been recently shown to combine
strong worst-case regression guarantees with elegant statistical
properties \citep{unbiased-estimates}.
However, these results are limited to the over-determined setting \citep{leveraged-volume-sampling,correcting-bias,minimax-experimental-design} and
ridge regression
\citep{regularized-volume-sampling,bayesian-experimental-design}.
Our results are also related to recent work on using DPPs
to analyze the expectation of the
inverse~\citep{determinantal-averaging,debiasing-second-order} and generalized
inverse~\citep{randomized-newton,nystrom-multiple-descent} of a subsampled matrix.

% Also in the over-determined setting, 
% \cite{correcting-bias-journal} provided evidence for the fact that
% determinantal rescaling can be used to modify the original data
% distribution (particularly, a multivariate Gaussian) without a
% significant distortion to the estimator, while making certain
% statistical quantities expressible analytically. We take this
% direction further by analyzing the unregularized least squares
% estimator in the under-determined setting which is less well
% understood, partly due to the presence of implicit regularization.

\textbf{Implicit regularization.}
The term implicit regularization typically refers to the notion that approximate
computation % (e.g., rather than exactly minimizing a function $f$,
% instead running an approximation algorithm to get an approximately
% optimal solution)
can implicitly lead to
statistical regularization. % (e.g., exactly minimizing an objective of
% the form $f + \lambda g$, for some well-specified $\lambda$ and $g$)
See~\cite{MO11-implementing, PM11, GM14_ICML} and references therein
for early work on the topic; and see~\cite{Mah12} for an overview.
More recently, often motivated by neural networks, there has been work
on implicit regularization that typically considered SGD-based
optimization algorithms.
See, e.g., theoretical results~\citep{NTS14_TR,Ney17_TR,SHNx17_TR,GWBNx17,ACHL19,KBMM19_TR}
as well as extensive empirical studies~\citep{MM18_TR,MM19_HTSR_ICML}.
The implicit regularization observed by us is different in that it is
not caused by an inexact approximation algorithm (such as SGD) but rather by the
selection of one out of many exact solutions (e.g., the minimum norm
solution). In this context, most relevant are the
asymptotic results of \cite{KLS18_TR} and \cite{LJB19_TR}.

\section{Surrogate random designs}
\label{s:determinantal}

In this section, we provide the definition of our surrogate random
design $S_\mu^n$, where $\mu$ is a $d$-variate probability measure and
$n$ is the sample size. This distribution is used in place
of the standard random design $\mu^n$ consisting of $n$ row vectors drawn
independently from $\mu$.
% Our surrogate design uses determinantal
% rescaling to alter the joint distribution of the vectors so that
% certain expected quantities (such as the mean squared error of the
% Moore-Penrose estimator) can be expressed in a closed form.
% We start by introducing notation.
%to be used throughout the rest of the paper.

%\paragraph{Preliminaries}
\textbf{Preliminaries.}
%The set $\{1,...,n\}$ will be denoted by $[n]$.
For an $n\times n$ matrix $\A$, we use $\pdet(\A)$ to denote the pseudo-determinant of $\A$,
which is the product of non-zero eigenvalues (repeated eigenvalues are taken to the power of their
algebraic multiplicity).
For index subsets $\Ic$
and $\Jc$, we use $\A_{\Ic,\Jc}$ to denote the submatrix of $\A$ with
rows indexed by $\Ic$ and columns indexed by $\Jc$. We may write
$\A_{\Ic,*}$ to indicate that we take a subset of rows.
% We use $\adj(\A)$ to denote the adjugate of $\A$, defined as follows: the
% $(i,j)$th entry of $\adj(\A)$ is
% $(-1)^{i+j}\det(\A_{[n]\backslash\{j\},[n]\backslash\{i\}})$.
% We will
% use two useful identities related to the adjugate: (1)
% $\adj(\A)=\det(\A)\A^{-1}$ for invertible $\A$, and (2)
% $\det(\A+\u\v^\top)=\det(\A)+\v^\top\!\adj(\A)\u$.
% A formula called Sylvester's theorem
% relates the adjugate and the determinant: $\det(\A+\u\v^\top)=\det(\A)+\v^\top\!\adj(\A)\u$.
% For a probability measure $\mu$ over $\R^d$, we use $\x^\top\!\sim\mu$
% to denote a random row vector $\x^\top$ sampled according to this distibution.
We let $\X\sim\mu^k$ denote a $k\times d$ random matrix with rows
drawn i.i.d.~according to $\mu$, and the $i$th row is denoted as $\x_i^\top$.
We also let $\Sigmab_\mu=\E_{\mu}[\x\x^\top]$, where $\E_{\mu}$ refers to
the expectation with respect to $\x^\top\!\sim\mu$, assuming throughout that
$\Sigmab_\mu$ is well-defined and positive definite.
We use $\Poisson(\gamma)_{\leq a}$ as the Poisson distribution
restricted to $[0,a]$, whereas $\Poisson(\gamma)_{\geq a}$ is restricted
to $[a,\infty)$.
We also let $\#(\X)$ denote the number of rows of $\X$.

% We now define a family of determinantal distributions over random matrices $\Xb$,
% where not only the entries but also the number of rows is
% randomized.
% This randomized sample size is a crucial property of our
% designs that enables our analysis.
% Our definition follows by
% expressing $\E[F(\Xb)]$ for real-valued functions $F:\bigcup_{k\geq 0}\R^{k\times
%   d}\rightarrow \R$ (the expectation may be undefined for some functions).
\begin{definition}\label{d:det}
  Let $\mu$ satisfy Assumption~\ref{a:general-position} and let $K$ be
  a random variable over $\mathbb{Z}_{\geq 0}$. A determinantal design
  $\Xb\sim \Det(\mu,K)$ is a
  distribution with the same domain as $\X\sim\mu^K$ such that~for any
  event $E$ measurable w.r.t.~$\X$, we have\vspace{-2mm}
  \begin{align*}
    \Pr\big\{\Xb\in E\big\}\ = \frac{\E[\pdet(\X\X^\top)\one_{[\X\in E]}]}{\E[\pdet(\X\X^\top)]}.
  \end{align*}
\end{definition}\vspace{-2mm}

\noindent
% Setting $F(\cdot)$ to 1, observe that the proportionality constant
% must be $1/\E[\pdet(\X\X^\top)]$.
The above definition can be
interpreted as rescaling the density function of $\mu^K$ by the
pseudo-determinant, and then renormalizing it.
We now construct our surrogate design $S_\mu^n$ by appropriately
selecting the random variable $K$.
The obvious choice of $K=n$ does \emph{not} result in simple closed
form expressions for the MSE in the under-determined regime (i.e.,
$n<d$), which is the regime of primary interest to us.
Instead, we derive our random variables $K$ from the Poisson
distribution.
\begin{definition}\label{d:surrogate}
  For $\mu$ satisfying Assumption~\ref{a:general-position},
  define surrogate design $S_\mu^n$ as $\Det(\mu,K)$ where:
  \vspace{-2mm}
  \begin{enumerate}
    \item if $n<d$, then $K\sim \Poisson(\gamma_n)_{\leq d}$ with
          $\gamma_n$ as the solution of
          $n=\tr(\Sigmab_\mu(\Sigmab_\mu+\frac1{\gamma_n}\I)^{-1})$,
          \vspace{-2mm}
    \item if $n=d$, then we simply let $K=d$,
          \vspace{-2mm}
    \item if $n>d$, then $K\sim\Poisson(\gamma_n)_{\geq d}$ with $\gamma_n=n-d$.
  \end{enumerate}
\end{definition}

\noindent
Note that the under-determined case, i.e., $n<d$, is restricted to $K\leq d$ so that, under Assumption~\ref{a:general-position}, $\pdet(\X\X^\top)=\det(\X\X^\top)$ with probability 1.
On the other hand in the over-determined case, i.e., $n>d$, we have
$K\geq d$ so that $\pdet(\X\X^\top)=\det(\X^\top\X)$. In the special case
of $n=d=K$ both of these equations are satisfied: $\pdet(\X\X^\top)=\det(\X^\top\X)=\det(\X\X^\top)=\det(\X)^2$.

The first non-trivial property of the surrogate design $S_\mu^n$ is
that the expected sample size is in fact always equal to $n$, which we
prove in Appendix \ref{appx: proof-of-l-size}.
\begin{lemma} \label{l:size}
  Let $\Xb\sim S_\mu^n$ for any $n>0$.
  Then, we have $\E[\#(\Xb)] = n$.
\end{lemma}

\noindent
Our general template for computing expectations under
a surrogate design $\Xb\sim\S_\mu^n$ is to use the following expressions based on the
i.i.d.~random design $\X\sim\mu^K$:
\begin{align}
  \E[F(\Xb)] & =\begin{cases}
    \frac{\E[\det(\X\X^\top)F(\X)]}{\E[\det(\X\X^\top)]}
    \quad K\sim\Poisson(\gamma_n) & \text{for }n<d, \\[2mm]
    \frac{\E[\det(\X)^2F(\X)]}{\E[\det(\X)^2]}\hspace{3mm}
    \quad K=d                     & \text{for }n=d, \\[2mm]
    \frac{\E[\det(\X^\top\X)F(\X)]}{\E[\det(\X^\top\X)]}
    \quad K\sim\Poisson(\gamma_n) & \text{for }n>d.
  \end{cases}\label{eq:cases}
\end{align}
These formulas follow from Definitions \ref{d:det} and
\ref{d:surrogate} because the determinants $\det(\X\X^\top)$ and
$\det(\X^\top\X)$ are non-zero precisely in the regimes $n\leq d$ and
$n\geq d$, respectively, which is why we can drop the restrictions on the
range of the Poisson distribution.
We compute the normalization constants
by introducing the concept of determinant preserving random matrices,
discussed in Section \ref{s:dp}.
% can be obtained using the following formulas: if $\X\sim\mu^K$ then
% \begin{align*}
% \text{(Lemma~\ref{l:normalization})}&&
% \E\big[\det(\X\X^\top)\big]  &= \ee^{-\gamma_n}\det(\I +
% \gamma_n\Sigmab_\mu)&&\text{for } K\sim\Poisson(\gamma_n),\ n<d,\\
%   \text{(Lemma~\ref{l:cb})}&& \E\big[\det(\X)^2\big] &=
%   d!\det(\Sigmab_\mu),&&\text{for } K=n=d,\\
%   \text{(Lemma~\ref{l:poisson})}&& \E\big[\det(\X^\top\X)\big]
% &=\det(\gamma_n\Sigmab_\mu),&&  \text{for }K\sim\Poisson(\gamma_n),\ n>d.
% \end{align*}
% \begin{remark}
% We will use $Z_\mu^n$ as a shorthand for the above normalization constants.
% \end{remark}
% For the case of $n=d$, the normalization constant can be found in the
% literature \cite{expected-generalized-variance,correcting-bias}:
% We prove Lemmas \ref{l:poisson} and \ref{l:normalization} in Section
% \ref{s:dp} by introducing the
% concept of determinant preserving random matrices.
% The lemmas play a
% crucial role in deriving a number of new expectation formulas for the
% under- and over-determined surrogate designs that we use to prove Theorems~\ref{t:mse}
% and~\ref{t:unbiased} in Section~\ref{s:expectations}.
% On the other hand, Lemma  \ref{l:cb} and the design $S_\mu^d$ can be found in the
% literature \citep{expected-generalized-variance,correcting-bias-journal},
% and we will rely on those known results in this case.
% Importantly, the $n=d$ case offers a continuous transition between the under-
% and over-determined regimes because the distribution $S_\mu^n$
% converges to $S_\mu^d$ when $n$ approaches $d$ from above and
% below.

\paragraph{Proof sketch of Theorem \ref{t:mse}}
% We now highlight some key expectation formulas for surrogate designs, which are used to
% derive the MSE expressions from Theorem \ref{t:mse}. 
We focus here on the under-determined regime (i.e., $n<d$),
highlighting the key new expectation formulas we develop to derive the
MSE expressions for surrogate designs. A standard decomposition of the MSE yields:
\begin{align}
  \MSE{\Xb^\dagger\ybb}
   & = \E\big[\|\Xb^\dagger(\Xb\w^*+\xib)-\w^*\|^2\big]
  =\sigma^2\E\big[\tr\big((\Xb^\top\Xb)^{\dagger}\big)\big] +
  \w^{*\top}\E\big[\I-\Xb^\dagger\Xb\big]\w^*.\label{eq:mse-derivation}
\end{align}
Thus, our task is to find closed form expressions for the two
expectations above. The latter, which is the expected projection onto
the complement of the row-span of $\Xb$, is proven in
Appendix \ref{s:unbiased-proof}.
\begin{lemma}\label{l:proj}
  If  $\Xb\sim S_\mu^n$ and $n<d$, then we have:
  $\E\big[\I-\Xb^\dagger\Xb\big] = (\gamma_n\Sigmab_\mu+\I)^{-1}$.
\end{lemma}

\noindent
No such expectation formula is known for i.i.d.~designs, except when
$\mu$ is an isotropic Gaussian. In Appendix \ref{s:unbiased-proof}, we
also prove a generalization of Lemma \ref{l:proj} which is then used
to establish our implicit regularization result
(Theorem~\ref{t:unbiased}). We next give an expectation formula
for the trace of the Moore-Penrose inverse of the covariance
matrix for a surrogate design (proof in Appendix \ref{a:mse-proof}).
\begin{lemma}\label{l:sqinv-all}
  If  $\Xb\sim S_\mu^n$ and $n<d$, then:
  $\E\big[\tr\big((\Xb^\top\Xb)^{\dagger}\big)\big]
    =\gamma_n\big(1-
    \det\!\big((\tfrac1{\gamma_n}\I+\Sigmab_\mu)^{-1}\Sigmab_\mu\big)\big)$.
\end{lemma}

\noindent
% The case of $n=d$ in the above lemma was shown by
% \cite{correcting-bias-journal} in Theorem~2.12, whereas the case of $n>d$
% is closely related to their Theorem 2.9.
Note the implicit regularization term which appears in both formulas,
given by $\lambda_n=\frac1{\gamma_n}$. Since $n =
  \tr(\Sigmab_\mu(\Sigmab_\mu+\lambda_n\I)^{-1})=d-\lambda_n\tr((\Sigmab_\mu+\lambda_n\I)^{-1})$,
it follows that $\lambda_n=(d-n)/\tr((\Sigmab_\mu+\lambda_n\I)^{-1})$.
Combining this with Lemmas \ref{l:proj} and \ref{l:sqinv-all}, we
recover the surrogate MSE expression in Theorem~\ref{t:mse}.

\section{Determinant preserving random matrices}
\label{s:dp}

In this section, we introduce the key tool for computing expectation formulas of matrix determinants.
It is used in our analysis of the surrogate design, and it should be of independent interest.

The key question motivating the following definition is: \textit{When does taking expectation commute with computing a determinant for a square random matrix?}
\begin{definition}\label{d:main}
  A random $d\times d$ matrix $\A$ is called determinant
  preserving (d.p.), if
  \begin{align*}
    \E\big[\!\det(\A_{\Ic,\Jc})\big] =
    \det\!\big(\E[\A_{\Ic,\Jc}]\big)\quad \text{for all }\Ic,\Jc\subseteq
    [d]\text{ s.t. }|\Ic|=|\Jc|.
  \end{align*}
\end{definition}

\noindent
We next give a few simple examples to provide some intuition. First, note
that every $1\times 1$ random matrix is determinant preserving simply
because taking a determinant is an identity transfomation in one
dimension. Similarly, every fixed matrix is determinant preserving because
in this case taking the expectation is an identity
transformation. In all other cases, however, Definition \ref{d:main}
has to be verified more carefully. Further examples (positive and
negative) follow.
\begin{example}
  If $\A$ has i.i.d. Gaussian entries $a_{ij}\sim\Nc(0,1)$, then
  $\A$ is d.p.~because $\E[\det(\A)]=0$.
\end{example}

\noindent
In fact, it can be shown that all random matrices with independent entries
are determinant preserving. However, this is not a necessary condition.
\begin{example}\label{e:rank-1}
  Let $\A = s\,\Z$, where $\Z$ is fixed with $\rank(\Z) = r$, and $s$
  is a scalar random variable. Then for $|\Ic|=|\Jc|=r$ we have
  \begin{align*}
    \E\big[\det(s\,\Z_{\Ic,\Jc})\big] & = \E[s^r]\det(\Z_{\Ic,\Jc})
    =\det\Big(\big(\E[s^r]\big)^{\frac1r}\,\Z_{\Ic,\Jc}\Big),
  \end{align*}
  so if $r=1$ then $\A$ is determinant preserving, whereas if $r>1$
  and $\Var[s]>0$ then it is not.
\end{example}

\noindent
To construct more complex examples, we show that determinant preserving random matrices are
closed under addition and multiplication. The proof of this result is
an extension of an existing argument, given by
\cite{determinantal-averaging} in the proof of Lemma~7, for computing
the expected determinant of the sum of rank-1 random matrices (proof in Appendix \ref{a:dp}).
\begin{lemma}[Closure properties]\label{t:ring}
  If $\A$ and $\B$ are independent and determinant preserving, then:
  \vspace{-1mm}
  \begin{enumerate}
    \item $\A+\B$ is determinant preserving,
          \vspace{-2mm}
    \item $\A\B$ is determinant preserving.
  \end{enumerate}
\end{lemma}
% \begin{lemma}\label{t:ring}
%   If $\A,\B$ are independent and d.p.~then
%   $\A+\B$ and $\A\B$ are also determinant preserving.
% \end{lemma}

\noindent
Next, we introduce another important class of d.p.~matrices:
a sum of i.i.d.~rank-1 random matrices with the number of
i.i.d.~samples being a Poisson random variable. Our use of the Poisson
distribution is crucial for the below result to hold. It is an
extension of an expectation formula given by \cite{dpp-intermediate}
for sampling from discrete distributions (proof in Appendix \ref{a:dp}).
\begin{lemma}\label{l:poisson}
  If $K$ is a Poisson random variable and $\A,\B$ are random $K\times d$
  matrices whose rows  are sampled as an i.i.d.~sequence of joint pairs of
  random vectors, then $\A^\top\B$ is d.p., and so:
  \begin{align*}
    \E\big[\det(\A^\top\B)\big] & = \det\!\big(\E[\A^\top\B]\big).
  \end{align*}
\end{lemma}

\noindent
Finally, we show the expectation formula needed for obtaining the
normalization constant of the under-determined surrogate design, given
in \eqref{eq:cases}.
The below result is more general than the normalization constant
requires, because it allows the matrices $\A$ and $\B$ to be different
(the constant is obtained by setting $\A=\B=\X\sim\mu^K$).
In fact, we use this more general statement to show Theorems
\ref{t:mse} and~\ref{t:unbiased}. The proof uses
Lemmas \ref{t:ring} and \ref{l:poisson} (see Appendix \ref{a:dp}).
\begin{lemma}\label{l:normalization}
  If $K$ is a Poisson random variable and $\A$, $\B$ are random $K\times d$
  matrices whose rows  are sampled as an i.i.d.~sequence of joint pairs of
  random vectors, then
  \begin{align*}
    \E\big[\det(\A\B^\top)\big] & = \ee^{-\E[K]}\det\!\big(\I + \E[\B^\top\A]\big).
  \end{align*}
\end{lemma}

\input{continuous-empirical.tex}

\section{Conclusions}
\label{s:conclusions}

We derived exact non-asymptotic expressions for the MSE of the
Moore-Penrose estimator in the linear regression task, reproducing
the double descent phenomenon as the sample size crosses between the
under- and over-determined regime. To achieve this, we modified the
standard i.i.d.~random design distribution using a determinantal
point process to obtain a surrogate design which admits exact MSE expressions,
while capturing the key properties of the i.i.d.~design. We
also provided a result that relates the expected value of the
Moore-Penrose estimator of a training sample in the under-determined regime (i.e., the
minimum norm solution) to the ridge-regularized least squares solution
for the population distribution, thereby providing an interpretation for the
implicit regularization resulting from over-parameterization.

% Our work opens up a number of new directions for future research.
% This includes extending our surrogate analysis to the
% misspecified linear regression discussed above.
% Also, it remains open whether the analysis we provided for the mean
% squared error can be reproduced in the context of mean squared
% \emph{prediction} error, which is relevant in many machine learning
% tasks. Finally, while Theorem \ref{t:asymptotic} states that our surrogate
% expressions for the MSE under certain Gaussian designs are
% asymptotically consistent with the multiplicative error rate of
% $O(1/d)$, we believe that this fact extends to
% the setting not covered by the theorem: under-determined regime (i.e.,
% $n<d$) with a non-isotropic Gaussian distribution (i.e., $\Sigmab\neq
% \I$). We break down our analysis into verifying two
% conjectures which are of independent interest to multivariate
% Gaussian analysis.
\ifisarxiv\else
  \section*{Broader Impact}

  While the double descent phenomenon has been empirically observed in a variety
  of applications, mathematical descriptions of it are oftentimes complex and
  inaccesible to non-experts. In contrast, our surrogate design and the
  accompanying closed-form expressions provide an easily computable rule of thumb
  for estimating the generalization error. One important application is
  the high-dimensional (i.e. underdetermined, $n < d$) regime
  experienced by modern machine learning systems where the number of
  parameters vastly exceeds the quantity of
  available data. Our research can be applied here to provide a theoretical
  understanding of the surprising phenomenon where without proper
  regularization \citep{nakkiran2020optimal} more data (i.e. increasing $n$)
  may lead to worse generalization performance.

  Better theoretical understanding of generalization error in the small data
  regime has important societal impact. Through theoretically modeling a system's
  performance, we can build safer systems by better understanding how badly an
  estimator fails and accounting for these failure modes in the system's design.
  Furthermore, the improved understanding of minimum norm solutions performing
  worse with more data offers an appealing trade-off where certain systems can
  both improve their performance and respect the privacy of its users by
  collecting less data.
\fi

\paragraph{Acknowledgements.}
We would like to acknowledge ARO, DARPA, NSF, ONR, and GFSD for providing
partial support of this work. We also thank Zhenyu Liao for pointing
out fruitful connections between our results and the asymptotic
analysis of random matrix resolvents.

\ifisarxiv
  \bibliographystyle{plainnat}
\else
  \bibliographystyle{alpha}
\fi

\bibliography{../pap}

\ifisarxiv\else\newpage\fi

\appendix

\input{appendix}

\end{document}