### 2022-08-23 design/continuous_topml/appendix.tex

\section{Proof of Lemma~\ref{l:size}}
\label{appx: proof-of-l-size}

We first record an important property of the design $S_\mu^d$
which can be used to construct an over-determined design for any $n>d$. A similar
version of this result was also previously shown by
\cite{correcting-bias-journal} for a different determinantal design.

\begin{lemma}\label{l:decomposition}
  Let $\Xb\sim S_\mu^d$ and $\X\sim \mu^K$, where
  $K\sim\Poisson(\gamma)$. Then the matrix composed of a random
  permutation of the rows from $\Xb$ and $\X$ is distributed according to
  $S_\mu^{d+\gamma}$.
\end{lemma}

\begin{proof}
Let $\Xt$ denote the matrix constructed from the permuted rows of
$\Xb$ and $\X$.  Letting $\Z\sim\mu^{K+d}$, we derive the probability
$\Pr\big\{\Xt\!\in\! E\big\}$ by summing over the possible index subsets  $S\subseteq
[K+d]$ that correspond to the rows coming from $\Xb$:
\begin{align*}
  \Pr\big\{\Xt\in E\big\} &= \E\bigg[\frac{1}{\binom{K+d}{d}}
  \sum_{S:\,|S|=d}\frac{\E[\det(\Z_{S,*})^2\one_{[\Z\in E]}\mid
  K]}{d!\det(\Sigmab_\mu)}\bigg]\\
  &=\sum_{k=0}^\infty
    \frac{\gamma^k\ee^{-\gamma}}{k!}\,\frac{\gamma^dk!}{(k+d)!}\,
    \frac{\E\big[\sum_{S:\,|S|=d}\det(\Z_{S,*})^2\one_{[\Z\in E]}\mid
    K=k\big]}{\det(\gamma\Sigmab_\mu)}\\
  &\overset{(*)}{=} \sum_{k=0}^\infty
    \frac{\gamma^{k+d}\ee^{-\gamma}}{(k+d)!}
    \,\frac{\E[\det(\Z^\top\Z)\one_{[\Z\in E]}\mid K=k]}{\det(\gamma\Sigmab_\mu)},
\end{align*}
where $(*)$ uses the Cauchy-Binet formula to sum over all subsets $S$
of size $d$. Finally, since the sum shifts from $k$
to $k+d$, the last expression can be rewritten as
$\E[\det(\X^\top\X)\one_{[\X\in E]}]/\det(\gamma\Sigmab_\mu)$, where recall that
$\X\sim\mu^K$ and $K\sim\Poisson(\gamma)$, matching the definition of $S_\mu^{d+\gamma}$.
\end{proof}

We now proceed with the proof of Lemma \ref{l:size}, where we establish
that the expected sample size of $S_\mu^n$ is indeed $n$.

\begin{proofof}{Lemma}{\ref{l:size}}
  The result is obvious when $n=d$, whereas
  for $n>d$ it is an immediate consequence
  of Lemma \ref{l:decomposition}.
  Finally, for $n<d$ the expected sample
  size follows as a corollary of Lemma \ref{l:proj}, which states that
  \begin{align*}
\text{(Lemma \ref{l:proj})} \qquad\E\big[\I - \Xb^\dagger\Xb\big] =
    (\gamma_n\Sigmab_\mu + \I)^{-1},
  \end{align*}
  where $\Xb^\dagger\Xb$ is the orthogonal projection onto
  the subspace spanned by the rows of $\Xb$. Since the rank of this
  subspace is equal to the number of the rows, we have
  $\#(\Xb)=\tr(\Xb^\dagger\Xb)$, so
  \begin{align*}
    \E\big[\#(\Xb)\big] = d - \tr\big((\gamma_n\Sigmab_\mu +
    \I)^{-1}\big) =
    \tr\big(\gamma_n\Sigmab_\mu(\gamma_n\Sigmab_\mu+\I)^{-1}\big) = n,
  \end{align*}
  which completes the proof.
\end{proofof}

\section{Proofs for Section \ref{s:dp}}
\label{a:dp}

We use $\adj(\A)$ to denote the adjugate of $\A$, defined as follows: the
$(i,j)$th entry of $\adj(\A)$ is
$(-1)^{i+j}\det(\A_{[n]\backslash\{j\},[n]\backslash\{i\}})$.
We will use two useful identities related to the adjugate: (1)
$\adj(\A)=\det(\A)\A^{-1}$ for invertible $\A$, and (2)
$\det(\A+\u\v^\top)=\det(\A)+\v^\top\!\adj(\A)\u$
\citep[see Fact 2.14.2 in][]{matrix-mathematics}.

First, note that from the definition of an adjugate matrix it immediately follows that if $\A$ is
determinant preserving then adjugate commutes with expectation for this matrix:
\begin{align}
  \E\big[\big(\!\adj(\A)\big)_{i,j}\big] &=
  \E\big[(-1)^{i+j}\det(\A_{[d]\backslash\{j\},[d]\backslash\{i\}})\big]\nonumber
  \\
&=(-1)^{i+j}\det\!\big(\E[\A_{[d]\backslash\{j\},[d]\backslash\{i\}}]\big)
  \\
  &= \big(\!\adj(\E[\A])\big)_{i,j}.\label{eq:adj}
\end{align}
% The adjugate is useful in our analysis because it connects the
% determinant and the inverse via the formula
% $\adj(\A)=\det(\A)\A^{-1}$, which holds for any invertible $\A$.

\begin{proofof}{Lemma}{\ref{t:ring}} \
 First, we show that $\A+\u\v^\top$ is d.p.~for any fixed
 $\u,\v\in\R^d$. Below, we use the identity for a rank one
 update of a determinant:
 $\det(\A+\u\v^\top)=\det(\A)+\v^\top\!\adj(\A)\u$. It follows that
 for any $\Ic$ and $\Jc$ of the same size,
  \begin{align*}
\E\big[\!\det(\A_{\Ic,\Jc}\!+\u_{\Ic}\v_{\Jc}^\top)\big] &=
    \E\big[\!\det(\A_{\Ic,\Jc}) +
    \v_{\Jc}^\top\adj(\A_{\Ic,\Jc}) \u_{\Ic}\big]\\
    &\overset{(*)}{=}\det\!\big(\E[\A_{\Ic,\Jc}]\big) +
      \v_{\Jc}^\top\adj\!\big(\E[\A_{\Ic,\Jc}]\big) \u_{\Ic}\\
    &=\det\!\big(\E[\A_{\Ic,\Jc} \!+ \u_{\Ic}\v_{\Jc}^\top]\big),
  \end{align*}
  where $(*)$ used \eqref{eq:adj}, i.e., the fact that for d.p.~matrices, adjugate commutes
  with expectation. Crucially, through the definition of an adjugate
  this step implicitly relies on the assumption that all the square
  submatrices of $\A_{\Ic,\Jc}$ are also  determinant preserving.
  Iterating this, we get that $\A+\Z$ is d.p.~for any fixed
  $\Z$. We now show the same for $\A+\B$:
  \begin{align*}
\E\big[\!\det(\A_{\Ic,\Jc}\!+\B_{\Ic,\Jc})\big]
    &=
      \E\Big[\E\big[\!\det(\A_{\Ic,\Jc}\!+\B_{\Ic,\Jc})\mid\B\big]\Big]\\
    &\overset{(*)}{=}\E\Big[\!\det\!\big(\E[\A_{\Ic,\Jc}]\!+\B_{\Ic,\Jc}\big)\Big]\\
      &= \det\!\big(\E[\A_{\Ic,\Jc}\!+\B_{\Ic,\Jc}]\big),
  \end{align*}
  where $(*)$  uses the fact that after conditioning on $\B$ we can
  treat it as a fixed matrix. Next, we show that $\A\B$ is determinant preserving via the Cauchy-Binet formula:
  \begin{align*}
    \E\big[\!\det\!\big((\A\B)_{\Ic,\Jc}\big)\big]
    &= \E\big[\!\det(\A_{\Ic,*}\B_{*,\Jc})\big]\\
    &=\E\bigg[\sum_{S:\,|S|=|\Ic|}\!\!\det\!\big(\A_{\Ic,S}\big)
      \det\!\big(\B_{S,\Jc}\big)\bigg]\\
&=\!\!\sum_{S:\,|S|=|\Ic|}\!\!\det\!\big(\E[\A]_{\Ic,S}\big)
                                                \det\!\big(\E[\B]_{S,\Jc}\big)\\
    &=\det\!\big(\E[\A]_{\Ic,*}\, \E[\B]_{*,\Jc}\big)\\
      &= \det\!\big(\E[\A\B]_{\Ic,\Jc}\big),
  \end{align*}
  where recall that $\A_{\Ic,*}$ denotes the submatrix of $\A$
  consisting of its (entire) rows indexed by $\Ic$.
  \end{proofof}

To prove Lemma \ref{l:poisson}, we will use the following
lemma, many variants of which appeared in the literature
\cite[e.g.,][]{expected-generalized-variance}. We use the one given by
\cite{correcting-bias}.
\begin{lemma}[\cite{correcting-bias}]\label{l:cb}
If the rows of random $k\times d$ matrices $\A,\B$
  are sampled as an i.i.d.~sequence of $k\geq d$ pairs of joint random vectors, then
\begin{align}
  k^d\,\E \big[\det(\A^\top\B)\big]
  &= \ktd\,\det\!\big(\E[\A^\top\B]\big).
     \end{align}
 \end{lemma}

\noindent
Here, we use the following standard shorthand: $\ktd =
\frac{k!}{(k-d)!} = k\,(k-1)\dotsm(k-d+1)$. Note that the above result
almost looks like we are claiming that the matrix $\A^\top\B$ is d.p.,
but in fact it is not because $k^d\neq \ktd$. The difference
in those factors is precisely what we are going to correct with the
Poisson random variable. We now present the proof of Lemma
\ref{l:poisson}.
\begin{proofof}{Lemma}{\ref{l:poisson}}
Without loss of generality, it suffices to check Definition \ref{d:main} with both $\Ic$ and
$\Jc$ equal $[d]$. We first expand the expectation by
conditioning on the value of $K$ and letting $\gamma=\E[K]$:
    \begin{align*}
      \E\big[\!\det(\A^\top\B)\big]
      &= \sum_{k=0}^\infty
\E\big[\det(\A^\top\B)\mid K\!=\!k\big]\
\Pr(K\!=\!k)\\
      \text{(Lemma \ref{l:cb})}
      \quad&=
        \sum_{k=d}^\infty\frac{k! k^{-d}}{(k-d)!}\det\!\big(\E[\A^\top\B\mid
        K\!=\!k]\big)
        \frac{\gamma^k\ee^{-\gamma}}{k!}\\
      &=\sum_{k=d}^\infty
\Big(\frac\gamma k\Big)^d\det\!\big(\E[\A^\top\B\mid K\!=\!k]\big)
        \frac{\gamma^{k-d}\ee^{-\gamma}}{(k-d)!}.
     %  \\
     %  &=\sum_{k=0}^\infty \det\!\big(\E[\A^\top\B]\big)\,\Pr(K\!=\!k)
     % \ =\ \det\!\big(\E[\A^\top\B]\big).
    \end{align*}
    Note that $\frac\gamma k\,\E[\A^\top\B\mid K\!=\!k]=\E[\A^\top\B]$,
    which is independent of $k$. Thus we can rewrite the above
    expression as:
    \begin{align*}
\det\!\big(\E[\A^\top\B]\big)\sum_{k=d}^\infty\frac{\gamma^{k-d}\ee^{-\gamma}}{(k-d)!}
      =
      \det\!\big(\E[\A^\top\B]\big)\sum_{k=0}^\infty
      \frac{\gamma^{k}\ee^{-\gamma}}{k!}=\det\!\big(\E[\A^\top\B]\big),
    \end{align*}
    which concludes the proof.
  \end{proofof}

To prove Lemma \ref{l:normalization}, we use the following standard
determinantal formula which is used to derive the normalization
constant of a discrete determinantal point process.
\begin{lemma}[\cite{dpp-ml}]\label{l:det-standard}
  For any $k\times d$ matrices $\A,\B$ we have
  \[\det(\I+\A\B^\top)=\sum_{S\subseteq[k]}\det(\A_{S,*}\B_{S,*}^\top).\]
\end{lemma}

\begin{proofof}{Lemma}{\ref{l:normalization}}
By Lemma \ref{l:poisson}, the matrix $\B^\top\A$ is determinant
preserving. Applying Lemma \ref{t:ring} we conclude that
$\I+\B^\top\A$ is also d.p., so
\begin{align*}
  \det\!\big(\I+\E[\B^\top\A]\big) = \E\big[\det(\I+\B^\top\A)\big] =\E\big[\det(\I+\A\B^\top)\big],
\end{align*}
where the second equality is known as Sylvester's Theorem.
We rewrite the expectation of $\det(\I+\A\B^\top)$ by applying Lemma
\ref{l:det-standard}.  Letting $\gamma=\E[K]$, we obtain:
\begin{align*}
\E\big[\det(\I+\A\B^\top)\big]  &=\E\bigg[\sum_{S\subseteq [K]}\E\big[\det(\A_{S,*}\B_{S,*}^\top)\mid
    K\big]\bigg]\\
  &\overset{(*)}{=}\sum_{k=0}^\infty\frac{\gamma^k\ee^{-\gamma}}{k!}
  \sum_{i=0}^k\binom{k}{i} \E\big[\det(\A\B^\top)\mid K=i\big]\\
  &=\sum_{i=0}^\infty \E\big[\det(\A\B^\top)\mid K=i\big]
  \sum_{k\geq i}^\infty \binom{k}{i}
  \frac{\gamma^k\ee^{-\gamma}}{k!}\\
  &=\sum_{i=0}^\infty
    \frac{\gamma^i\ee^{-\gamma}}{i!}\E\big[\det(\A\B^\top)\mid K=i\big]
    \sum_{k\geq i}^\infty\frac{\gamma^{k-i}}{(k-i)!} = \E\big[\det(\A\B^\top)\big]\cdot\ee^\gamma,
\end{align*}
where $(*)$ follows from the exchangeability of the rows of $\A$ and
$\B$, which implies that the distribution of $\A_{S,*}\B_{S,*}^\top$ is the
same for all subsets $S$ of a fixed size $k$.
\end{proofof}

\section{Proof of Theorem \ref{t:mse}}
\label{a:mse-proof}
In this section we use $Z_\mu^n$ to denote the normalization
constant that appears in \eqref{eq:cases} when computing an expectation for surrogate design
$S_\mu^n$.
We first prove Lemma \ref{l:sqinv-all}. % by splitting the under- and
% over-determined cases and start with proving the former.
% Note that those
% results require $\mu$ to satisfy general position (Assumption \ref{a:general-position}),
% which implies that if $\X\sim\mu^k$ for $k\leq d$ then $\rank(\X)=k$.
\begin{lemma}[restated Lemma \ref{l:sqinv-all}]\label{l:sqinv-under}
If  $\Xb\sim S_\mu^n$ for $n<d$, then we have
\begin{align*}
    \E\big[\tr\big((\Xb^\top\Xb)^{\dagger}\big)\big]
    &={\gamma_n}\big(1- \det\!\big((\tfrac1{\gamma_n}\I+\Sigmab_\mu)^{-1}\Sigmab_\mu\big)\big).
\end{align*}
\end{lemma}
\begin{proof}
Let $\X\sim\mu^K$ for $K\sim\Poisson({\gamma_n})$. Note that if
$\det(\X\X^\top)>0$ then using the fact that
$\det(\A)\A^{-1}=\adj(\A)$ for any invertible matrix $\A$, we can write:
  \begin{align*}
    \det(\X\X^\top)\tr\big((\X^\top\X)^{\dagger}\big)
    &= \det(\X\X^\top)\tr\big((\X\X^\top)^{-1}\big) \\
    &= \tr(\adj(\X\X^\top)) \\[-1mm]
    &= \sum_{i=1}^K\det(\X_{-i}\X_{-i}^\top),
  \end{align*}
  where $\X_{-i}$ is a shorthand for $\X_{[K]\backslash\{i\},*}$.
Assumption \ref{a:general-position} ensures that
$\Pr\big\{\det(\X\X^\top)>0\big\}=1$, which allows us to write:
  \begin{align*}
Z_\mu^n\cdot \E\big[\tr\big((\Xb^\top\Xb)^{\dagger}\big)\big]
    &=\E\bigg[
    \sum_{i=1}^K\det(\X_{-i}\X_{-i}^\top)\ \big|\
    \det(\X\X^\top)>0\bigg]\cdot\overbrace{\Pr\big\{\det(\X\X^\top)>0\big\}}^{1}\\
    &=\sum_{k=0}^d\frac{\gamma_n^{k}\ee^{-\gamma_n}}{k!}\E\Big[
      \sum_{i=1}^k\det(\X_{-i}\X_{-i}^\top)\ \big|\  K=k\Big]\\
    &=\sum_{k=0}^d\frac{\gamma_n^{k}\ee^{-\gamma_n}}{k!}\, k\
      \E\big[\det(\X\X^\top)\mid K=k-1\big]\\
    &=\gamma_n\sum_{k=0}^{d-1}\frac{\gamma_n^{k}\ee^{-\gamma_n}}{k!}
      \E\big[\det(\X\X^\top)\mid K=k\big]\\
    &=\gamma_n\Big(\E\big[\det(\X\X^\top)\big]\ -\
      \frac{\gamma_n^{d}\ee^{-\gamma_n}}{d!}\E\big[\det(\X)^2\mid K=d\big]
      \Big) \\
    &\overset{(*)}{=}\gamma_n\big(\ee^{-\gamma_n}\det(\I +\gamma_n\Sigmab_\mu) -
      \ee^{-\gamma_n}\det(\gamma_n\Sigmab_\mu)\big),
  \end{align*}
  where $(*)$ uses Lemma \ref{l:normalization} for the first term and
  Lemma \ref{l:cb} for the second term. We obtain the desired result by
  dividing both sides by
  $Z_\mu^n=\ee^{-\gamma_n}\det(\I+\gamma_n\Sigmab_\mu)$.
\end{proof}
In the over-determined regime, a more general matrix expectation
formula can be shown (omitting the trace). The following result is
related to an expectation formula derived by
\cite{correcting-bias-journal}, however they use a slightly
different determinantal design so the results are incomparable.
\begin{lemma}\label{l:sqinv-over}
If $\Xb\sim S_\mu^n$ and $n>d$, then we
have
\begin{align*}
  \E\big[ (\Xb^\top\Xb)^{\dagger}\big] =
  \Sigmab_\mu^{-1}\cdot \frac{1-\ee^{-\gamma_n}}{\gamma_n}.
\end{align*}
\end{lemma}
\begin{proof}
Let $\X\sim\mu^K$ for $K\sim\Poisson(\gamma_n)$. Assumption
\ref{a:general-position} implies that for $K\neq d-1$ we have
\begin{align}
  \det(\X^\top\X)(\X^\top\X)^\dagger=\adj(\X^\top\X),\label{eq:adj-over}
  \end{align}
however when $k=d-1$ then \eqref{eq:adj-over} does not hold because
$\det(\X^\top\X)=0$ while $\adj(\X^\top\X)$ may be non-zero. It
follows that:
  \begin{align*}
Z_\mu^n\cdot
    \E\big[ (\Xb^\top\Xb)^{\dagger}\big]
    &=\E\big[\det(\X^\top\X)(\X^\top\X)^\dagger\big]\\
    &=\E\big[\adj(\X^\top\X)\big]-
\frac{\gamma_n^{d-1}\ee^{-\gamma_n}}{(d-1)!}
      \E\big[\adj(\X^\top\X)\mid K=d-1\big]\\
    &\overset{(*)}{=}\adj\!\big(\E[\X^\top\X]\big) -
      \frac{\gamma_n^{d-1}\ee^{-\gamma_n}}{(d-1)^{d-1}}
      \adj\!\big(\E[\X^\top\X\mid K=d-1]\big)\\
    &=\adj(\gamma_n\Sigmab_\mu) - \ee^{-\gamma_n}\adj(\gamma_n\Sigmab_\mu)\\
    &=\det(\gamma_n\Sigmab_\mu)\,(\gamma_n\Sigmab_\mu)^{-1}(1-\ee^{-\gamma_n})\\
    &=\det(\gamma_n\Sigmab_\mu)\,\Sigmab_\mu^{-1}\cdot\frac{1-\ee^{-\gamma_n}}{\gamma_n},
  \end{align*}
  where the first term in $(*)$ follows from Lemma
  \ref{l:normalization} and \eqref{eq:adj}, whereas the second term comes
  from Lemma 2.3 of \cite{correcting-bias-journal}.
Dividing both sides by $Z_\mu^n=\det(\gamma_n\Sigmab_\mu)$ completes the proof.
\end{proof}

Applying the closed form expressions from Lemmas
\ref{l:proj}, \ref{l:sqinv-all} and \ref{l:sqinv-over}, we derive
the formula for the MSE and prove Theorem \ref{t:mse} (we defer the
proof of Lemma \ref{l:proj} to Appendix \ref{s:unbiased-proof}).
\begin{proofof}{Theorem}{\ref{t:mse}}
  First, assume that $n<d$, in which case we have
  $\gamma_n=\frac1{\lambda_n}$ and moreover
  \begin{align*}
    n &= \tr\big(\Sigmab_\mu(\Sigmab_\mu+\lambda_n\I)^{-1}\big)\\
      &=\tr\big((\Sigmab_\mu+\lambda_n\I-\lambda_n\I)(\Sigmab_\mu+\lambda_n\I)^{-1}\big)\\
    &=d - \lambda_n\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big),
    %\frac{\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)}{d-n},
  \end{align*}
so we can write $\lambda_n$ as $(d-n)/\tr((\Sigmab_\mu+\lambda_n\I)^{-1})$.
  From this and Lemmas \ref{l:proj} and \ref{l:sqinv-under}, we
obtain the desired expression, where recall
  that $\alpha_n = \det\!\big(\Sigmab_\mu (\Sigmab_\mu+\frac1{\gamma_n})^{-1}\big)$:
  \begin{align*}
    \MSE{\Xb^\dagger\ybb} &= \sigma^2\,\gamma_n(1-\alpha_n) +
    \tfrac1{\gamma_n} \,\w^{*\top}(\Sigmab_\mu+\tfrac1{\gamma_n}\I)^{-1}\w^*
    \\
    &\overset{(a)}{=}\sigma^2\,\frac{1-\alpha_n}{\lambda_n} +
    \lambda_n\,\w^{*\top}(\Sigmab_\mu+\lambda_n\I)^{-1}\w^*\\
    &\overset{(b)}{=}\sigma^2\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)\frac{1-\alpha_n}{d-n}
      +
      (d-n)\frac{\w^{*\top}(\Sigmab_\mu+\lambda_n\I)^{-1}\w^*}
      {\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)}.
  \end{align*}
  While the expression given after $(a)$ is simpler than the one
after $(b)$, the latter better illustrates how the MSE depends on
the sample size $n$ and the dimension $d$.
  Now, assume that $n>d$. In this case, we have $\gamma_n=n-d$ and apply Lemma
  \ref{l:sqinv-over}:
  \begin{align*}
    \MSE{\Xb^\dagger\ybb}
    &= \sigma^2\,\tr(\Sigmab_\mu^{-1})\,
\frac{1-\ee^{-\gamma_n}}{\gamma_n}
=\sigma^2\,\tr(\Sigmab_\mu^{-1})
\,\frac{1-\beta_n}{n-d}.
  \end{align*}
The case of $n=d$ was shown in Theorem~2.12 of \cite{correcting-bias-journal}.
This concludes the proof.
\end{proofof}

\section{Proof of Theorem \ref{t:unbiased}}
\label{s:unbiased-proof}

As in the previous section, we use $Z_\mu^n$ to denote the normalization
constant that appears in \eqref{eq:cases} when computing an expectation
for surrogate design $S_\mu^n$.
Recall that our goal is to compute the expected value of
$\Xb^\dagger\ybb$ under the surrogate design $S_\mu^n$. Similarly as for Theorem
\ref{t:mse}, the case of $n=d$ was shown in Theorem 2.10 of
\cite{correcting-bias-journal}. We break the rest down into the
under-determined case $(n<d)$ and the over-determined case ($n>d$),
starting with the former. Recall that we do \emph{not} require any
modeling assumptions on the responses.
\begin{lemma}\label{l:ridge-under}
If $\Xb\sim S_\mu^n$ and $n<d$, then for any $y(\cdot)$
such that $\E_{\mu,y}[y(\x)\,\x]$ is well-defined,
denoting $\yb_i$ as $y(\xbb_i)$, we have
\begin{align*}
  \E\big[\Xb^\dagger \ybb\big]
  &=
    \big(\Sigmab_\mu+\tfrac1{\gamma_n}\I\big)^{-1}\E_{\mu,y}[y(\x)\,\x].
\end{align*}
\end{lemma}
\begin{proof}
   Let $\X\sim\mu^K$ for $K\sim\Poisson(\gamma_n)$ and denote
   $y(\x_i)$ as $y_i$.
  Note that when $\det(\X\X^\top)>0$, then
  the $j$th entry of $\X^\dagger\y$ equals
  $\f_j^\top(\X\X^\top)^{-1}\y$, where $\f_j$ is the $j$th
  column of $\X$, so:
\begin{align*}
  \det(\X\X^\top)\,(\X^\dagger\y)_j
  &= \det(\X\X^\top)\, \f_j^\top(\X\X^\top)^{-1}\y \\
  &=
  \det(\X\X^\top+\y\f_j^\top) - \det(\X\X^\top).
\end{align*}
If $\det(\X\X^\top)=0$, then also
$\det(\X\X^\top+\y\f_j^\top)=0$, so we can write:
\begin{align*}
Z_\mu^n\cdot\E\big[(\Xb^\dagger\ybb)_j\big]
  &=  \E\big[\det(\X\X^\top)(\X^\dagger\y)_j\big] \\
  &= \E\big[\det(\X\X^\top+\y\f_j^\top)-\det(\X\X^\top)\big]  \\
  &=\E\big[\det\!\big([\X,\y][\X,\f_j]^\top\big)\big] - \E\big[\det(\X\X^\top)\big]\\
  &\overset{(a)}{=}\ee^{-\gamma_n}\det\!\bigg(\I +
    \gamma_n\,\E_{\mu,y}\bigg[\begin{pmatrix}\x\x^\top&
      \!\!\x\, y(\x)\\ x_j\,\x^\top&\!\!
      x_j\,y(\x)\end{pmatrix}\bigg]\bigg)
                          -\ee^{-\gamma_n}\det(\I+\gamma_n\Sigmab_\mu)\\
  &\overset{(b)}{=}\ee^{-\gamma_n}\det(\I+\gamma_n\Sigmab_\mu) \\
    &\qquad \times \Big(\E_{\mu,y}\big[\gamma_n
    x_j\,y(\x)\big] - \E_{\mu}\big[\gamma_n
    x_j\,\x^\top\big](\I+\gamma_n\Sigmab_\mu)^{-1}\E_{\mu,y}\big[\gamma_n\x\,
    y(\x)\big]\Big),
\end{align*}
where $(a)$ uses Lemma \ref{l:normalization} twice, with the first
application involving two different matrices $\A=[\X,\y]$ and
$\B=[\X,\f_j]$, whereas $(b)$ is a standard determinantal identity
\cite[see Fact 2.14.2 in][]{matrix-mathematics}.
  Dividing both sides by $Z_\mu^n$ and letting $\v_{\mu,y}=\E_{\mu,y}[y(\x)\,\x]$, we obtain that:
  \begin{align*}
    \E\big[\Xb^\dagger\ybb\big]
    &= \gamma_n\v_{\mu,y} - \gamma_n^2\Sigmab_\mu(\I+\gamma_n\Sigmab_\mu)^{-1}\v_{\mu,y}\\
&=\gamma_n\big(\I - \gamma_n\Sigmab_\mu (\I+\gamma_n\Sigmab_\mu)^{-1}\big)\v_{\mu,y}
=\gamma_n(\I+\gamma_n\Sigmab_\mu)^{-1}\v_{\mu,y},
  \end{align*}
  which completes the proof.
\end{proof}
We return to Lemma \ref{l:proj}, regarding the expected orthogonal
projection onto the complement of the row-span of $\Xb$, i.e.,
$\E[\I-\Xb^\dagger\Xb]$, which follows as a corollary of Lemma~\ref{l:ridge-under}.
\begin{proofof}{Lemma}{\ref{l:proj}}
  We let $y(\x)=x_j$ where $j\in[d]$ and apply Lemma
  \ref{l:ridge-under} for each $j$, obtaining:
  \begin{align*}
    \I - \E\big[\Xb^\dagger\Xb] = \I -
    (\Sigmab_\mu+\tfrac1{\gamma_n}\I)^{-1}\Sigmab_\mu,
  \end{align*}
  from which the result follows by simple algebraic manipulation.
\end{proofof}

We move on to the over-determined case, where the ridge regularization
of adding the identity to $\Sigmab_\mu$ vanishes. Recall that we
assume throughout the paper that $\Sigmab_\mu$ is invertible.
\begin{lemma}\label{l:ridge-over}
  If $\Xb\sim S_\mu^n$ and $n>d$, then for any real-valued random function $y(\cdot)$
  such that $\E_{\mu,y}[y(\x)\,\x]$ is well-defined,
denoting $\yb_i$ as $y(\xbb_i)$, we have
 \begin{align*}
  \E\big[\Xb^\dagger \ybb\big]
  &=\Sigmab_\mu^{-1}\E_{\mu,y}\big[y(\x)\,\x\big].
\end{align*}
\end{lemma}
\begin{proof}
   Let $\X\sim\mu^K$ for $K\sim\Poisson(\gamma_n)$ and denote
   $y_i=y(\x_i)$. Similarly as in the proof of
   Lemma~\ref{l:ridge-under}, we note that when $\det(\X^\top\X)>0$,
   then
  the $j$th entry of $\X^\dagger\y$ equals
  $\e_j^\top(\X^\top\X)^{-1}\X^\top\y$, where $\e_j$ is the $j$th
standard basis vector, so:
\begin{align*}
  \det(\X^\top\X)\,(\X^\dagger\y)_j =
  \det(\X^\top\X)\, \e_j^\top(\X^\top\X)^{-1}\X^\top\y =
  \det(\X^\top\X+\X^\top\y\e_j^\top) - \det(\X^\top\X).
\end{align*}
If $\det(\X^\top\X)=0$, then also
$\det(\X^\top\X+\X^\top\y\e_j^\top)=0$. We proceed to compute the
expectation:
\begin{align*}
Z_\mu^n\cdot\E\big[(\Xb^\dagger\ybb)_j\big]
  &=  \E\big[\det(\X^\top\X)(\X^\dagger\y)_j\big] \\
  &= \E\big[\det(\X^\top\X+\X^\top\y\e_j^\top)-\det(\X^\top\X)\big]  \\
  &=\E\big[\det\!\big(\X^\top(\X+\y\e_j^\top)\big)\big] - \E\big[\det(\X^\top\X)\big]\\
  &\overset{(*)}{=}\det\!\Big(
    \gamma_n\,\E_{\mu,y}\big[\x(\x+ y(\x)\e_j)^\top\big]\Big)
    -\det(\gamma_n\Sigmab_\mu)\\
  &=\det\!\big(\gamma_n\Sigmab_\mu + \gamma_n\E_{\mu,y}[\x\,y(\x)]\e_j^\top\big)
    -\det(\gamma_n\Sigmab_\mu)\\
  &=\det(\gamma_n\Sigmab_\mu)\cdot
    \gamma_n\e_j^\top(\gamma_n\Sigmab_\mu)^{-1}\E_{\mu,y}\big[y(\x)\,\x\big],
\end{align*}
where $(*)$ uses Lemma \ref{l:poisson} twice (the first time, with
$\A=\X$ and $\B=\X+\y\e_j^\top$). Dividing both sides by
$Z_\mu^n=\det(\gamma_n\Sigmab_\mu)$ concludes the proof.
\end{proof}

\noindent
We combine Lemmas \ref{l:ridge-under} and \ref{l:ridge-over} to obtain
the proof of Theorem \ref{t:unbiased}.
\begin{proofof}{Theorem}{\ref{t:unbiased}}
The case of $n=d$ follows directly from Theorem~2.10 of
\cite{correcting-bias}.
Assume that $n<d$. Then we have
$\gamma_n=\frac1{\lambda_n}$, so the result follows
from Lemma \ref{l:ridge-under}.
% , noting that unlike in the lemma,
% Theorem \ref{t:unbiased} allows for the response model $y(\cdot)$ to be randomized:
% \begin{align*}
%   \E[\Xb^\dagger\ybb] = \E\big[\E[\Xb^\dagger\ybb\mid\ybb]\big] =
% % \w_{\lambda_n}^* = \argmin_\w\E_\mu\big[(\x^\top\w-y(\x))^2\big] +
% %   \lambda_n\|\w\|^2 =
% %   \big(\Sigmab_\mu+\lambda_n\I\big)^{-1}
% %   \E_{\mu}\big[y(\x)\,\x\big].
% \end{align*}
If $n>d$, then the result follows from Lemma \ref{l:ridge-over}.
\end{proofof}

\section{Proof of Theorem \ref{t:asymptotic}}
\label{sec:proof-of-t-asymptotic}

The proof of Theorem \ref{t:asymptotic} follows the standard decomposition of
MSE in Equation~\ref{eq:mse-derivation}, and in the process,
establishes consistency of the variance and bias terms
independently. To this end, we introduce the following two useful
lemmas that capture the limiting behavior of the 
variance and bias terms, respectively.

\begin{lemma}\label{c:wishart}
  Under the setting of Theorem~\ref{t:asymptotic}, we have, as $n,d \to \infty$
with $n/d \to \bar c \in (0, \infty) \setminus \{ 1 \}$ that
\begin{align}
%\bigg|\frac{\E\big[\tr((\X^\top \X)^\dagger)\big]}{\Vc(\Sigmab,n)} -1\bigg| \to 0 \qquad\text{for}\quad\Vc(\Sigmab,n)= \frac{1-\alpha_n}{\lambda_n}
\begin{cases}
  \E\big[\tr((\X^\top \X)^\dagger)\big] - (1-\alpha_n) \lambda_n^{-1} \to 0, & \text{for }\bar c < 1, \\
  \E\big[\tr((\X^\top \X)^\dagger)\big] - \frac{1-\beta_n}{n - d} \cdot \tr \Sigmab^{-1} \to 0, & \text{for }\bar c > 1
\end{cases}
\label{eq:wishart}
\end{align}
where $\lambda_n\geq 0$ is the unique solution to
$n=\tr(\Sigmab(\Sigmab+\lambda_n\I)^{-1})$,
$\alpha_n=\det(\Sigmab(\Sigmab+\lambda_n\I)^{-1})$,
and $\beta_n = e^{d-n}$.
\end{lemma}

\noindent

The second term in the MSE derivation \eqref{eq:mse-derivation}, $\E[\I -
\X^\dag \X]$, involves the expectation of a projection onto the orthogonal
complement of a sub-Gaussian general position sample $\X$.
This term is zero when $n > d$, and for $n < d$
we prove in \cref{sec:proof-of-c-projection} that the surrogate design's
bias $\mathcal{B}(\Sigmab, n)$ provides an asymptotically consistent approximation
to all of the eigenvectors and eigenvalues:

\begin{lemma}\label{c:projection}
Under the setting of Theorem~\ref{t:asymptotic}, for $\w \in \R^d$ of bounded Euclidean norm (i.e., $\| \w \| \le C'$ for all $d$), we have, as $n,d \to \infty$ with $n/d \to \bar c \in (0,1)$ that
\begin{align}
%\sup_{\w\in\R^d\backslash\{\zero\}}\bigg|\frac{\w^\top\E[\I-\X^\dagger\X]\w}{\w^\top \Bc(\Sigmab,n)\w} - 1\bigg| \to 0 \qquad\text{for}\quad\Bc(\Sigmab,n) = \lambda_n(\Sigmab+\lambda_n\I)^{-1}
  \w^\top\E[\I-\X^\dagger\X]\w - \lambda_n \w^\top (\Sigmab + \lambda_n \I)^{-1}\w \to 0
  \label{eq:projection}
\end{align}
while $\I - \X^\dagger \X = 0$ for $\bar c > 1$.
\end{lemma}

\subsection{Proof of \cref{c:wishart}}
\label{sec:proof-of-c-wishart}

\subsubsection{The $\bar c \in (0,1)$ case}

For $n < d$, we first establish (1) $\liminf_n \lambda_n > 0$ and (2) $\alpha_n \to 0$.
To prove (1), by hypothesis $\Sigmab \succeq c \I$ for all $d$. Since $\frac{n}{d} < 1$,
we have (by definition of $\lambda_n$) for some $\delta > 0$
\begin{align*}
  1 - \delta
  > \frac{n}{d}
  = \frac{1}{d} \tr(\Sigmab(\Sigmab + \lambda_n \I)^{-1})
  > \frac{c}{c + \lambda_n}
\end{align*}
Rearranging, we have $\lambda_n > \frac{\delta c}{1 - \delta} > 0$.
For (2), let $(\tau_i)_{i \in [d]}$ denote the eigenvalues of $\Sigmab$.
Since $1 - x \leq e^{-x}$ and $C \I \succeq \Sigmab \succeq c \I$ for all $d$,
\begin{align*}
  \alpha_n
  = \prod_{i=1}^d \frac{\tau_i}{\tau_i + \lambda_n}
  \leq \left(\frac{C}{C + \lambda_n}\right)^d
  = \left( 1 - \frac{\lambda_n}{C + \lambda_n}\right)^d
  \leq \exp\left(-d \frac{\lambda_n}{C + \lambda_n}\right)
\end{align*}
and since $\lambda_n > 0$ eventually as $d \to \infty$ we have
$\alpha_n \to 0$ so that $(1-\alpha_n) \lambda_n^{-1} - \lambda_n^{-1} \to 0$.

% (1) and (2) together imply $\mathcal{V}(\Sigmab, n) \to \lambda_n^{-1}$, and
% by Slutsky's theorem
% \begin{align*}
%   \lim_{n,d} \left\lvert
%   \frac{\E[\tr((\X^\top \X)^\dag)] - \mathcal{V}(\Sigmab, n)}{\mathcal{V}(\Sigmab, n)}
%   \right\rvert
%   = \lim_{n,d} \left\lvert
%   \frac{\E[\tr((\X^\top \X)^\dag)] - \lambda_n^{-1}}{\lambda_n^{-1}}
%   \right\rvert
% \end{align*}
% Since $\liminf_n \lambda_n > 0$ was shown in (1),
As a consequence of (2) and Slutsky's theorem, it suffices to show
$\tr(\X^\top \X)^\dag - \lambda_n^{-1} \overset{d}{\to} 0$ as $n,d \to \infty$.
To do this, we consider the limiting behavior of $\tr
(\X^\top \X)^\dagger/n = \tr (\X \X^\top)^\dagger/n$ as $n/d \to \bar c \in
(0,1)$, for $\X = \Z \Sigmab^{\frac12}$ with $\Z \in \mathbb R^{n \times d}$
having i.i.d.~zero mean, unit variance sub-Gaussian entries, i.e., the behavior
of
\begin{equation}\label{eq:limit1}
  \lim_{n,d \to \infty} \lim_{z \to 0^+} \frac1n \tr \left( \frac1n \X \X^\top + z \I_n \right)^{-1}
\end{equation}
by definition of the pseudo-inverse.

The proof comes in three steps: (i) for fixed $z > 0$, consider the limiting behavior of $ \delta(z) \equiv \tr (\X \X^\top/n + z \I_n)^{-1}/n$ as $n,d \to \infty$ and state
\begin{equation}\label{eq:limit2}
  \lim_{n,d \to \infty} \delta(z) - m(z) \to 0
\end{equation}
almost surely for some $m(z)$ to be defined; (ii) show that both $\delta(z)$ and its derivate $\delta'(z)$ are uniformly bounded (by some quantity independent of $z>0$) so that by Arzela-Ascoli theorem, $\delta(z)$ converges uniformly to its limit and we are allowed to take $z \to 0^+$ in \eqref{eq:limit2} and state
\begin{equation}\label{eq:limit3}
  \lim_{z \to 0^+} \lim_{n,d \to \infty} \delta(z) - \lim_{z \to 0^+} m(z) \to 0
\end{equation}
almost surely, given that the limit $\lim_{z \to 0^+} m(z) \equiv m(0)$ exists and eventually (iii) exchange the two limits in \eqref{eq:limit3} with Moore-Osgood theorem, to reach
\[
  \lim_{n,d \to \infty} \lim_{z \to 0^+} \frac1n \tr \left(\frac1n \X \X^\top + z \I_n \right)^{-1} - m(0) \to 0.
\]

Step (i) follows from \cite{silverstein1995empirical} that, we have, for $z > 0$ that
\[
  \delta(z) \equiv \frac1n \tr \left( \frac1n \X \X^\top  + z \I_n \right)^{-1}  - m(z) \to 0
\]
almost surely as $n,d \to \infty$, for $m(z)$ the unique positive solution to
\begin{equation}\label{eq:def-m}
  m(z) = \left( z + \frac1n \tr \Sigmab (\I + m(z) \Sigmab)^{-1} \right)^{-1}.
\end{equation}

For the above step (ii), we use the assumption $\Sigmab \succeq c \I \succ 0$
for all $d$ large, so that with $\X = \Z \Sigmab^{\frac12}$, we have for large enough $n,d$ that
\[
  \lambda_{\min} (\X \X^\top/n) \ge \lambda_{\min} (\Z \Z^\top/n) \lambda_{\min} (\Sigmab) \ge \frac{c}2 (\sqrt{\bar c} - 1 )^2
\]
almost surely, where we used Bai-Yin theorem \cite{bai1993limit}, which states that the minimum eigenvalue of $\Z \Z^\top/n$ is almost surely larger than $( \sqrt{\bar c} - 1 )^2/2$ for $n<d$ sufficiently large. Note that here the case $\bar c = 1$ is excluded.

Observe that
\[
  |\delta(z)| = \left| \frac1n \tr \left( \frac1n \X \X^\top  + z \I_n \right)^{-1} \right| \le \frac1{ \lambda_{\min} (\X \X^\top/n) }
\]
and similarly for its derivative, so that we are allowed to take the $z \to 0^+$ limit. Note that the existence of the $\lim_{z \to 0^+} m(z)$ for $m(z)$ defined in \eqref{eq:def-m} is well known, see for example \cite{ledoit2011eigenvectors}. Then, by Moore-Osgood theorem we finish step (iii) and by concluding that
\[
  \tr (\X^\top \X)^\dagger - m(0) \to 0
\]
for $m(0) = \lambda_n^{-1}$ the unique solution to $\lambda_n^{-1} = \left( \frac1n \tr \Sigmab (\I + \lambda_n^{-1} \Sigmab)^{-1} \right)^{-1}$, or equivalently, to
\[
  n = \tr \Sigmab (\Sigmab + \lambda_n \I)^{-1}
\]
as desired.

\subsubsection{The $\bar c \in (1, \infty)$ case}

First note that as $n, d \to \infty$ with $n > d$, we have $\beta_n = e^{d-n} \to 0$ and it
it suffices to show
\[
  \tr (\X^\top \X)^\dagger - \frac1{n-d} \tr \Sigmab^{-1} \to 0
\]
almost surely to conclude the proof.

In the $\bar c \in (1, \infty)$ case, it is more convenient to work on the following co-resolvent
\[
  \lim_{n,d \to \infty} \lim_{z \to 0^+} \frac1n \tr \left( \frac1n \X^\top \X + z \I_d \right)^{-1}
\]
where we recall $\X^\top \X = \Sigmab^{\frac12} \Z^\top \Z \Sigmab^{\frac12} \in \R^{d \times d}$ and following the same three-step procedure as in the $\bar c < 1$ case above. The only difference is in step (i) we need to assess the asymptotic behavior of $\delta \equiv \tr (\X^\top \X/n + z \I_d)^{-1}/n$. This was established in \cite{bai1998no} where it was shown that, for $z > 0$ we have
\[
  \frac1n \tr (\X^\top \X/n + z \I_d)^{-1} - \frac{d}n m(z) \to 0
\]
almost surely as $n,d \to \infty$, for $m(z)$ the unique solution to
\[
  m(z) = \frac1d \tr \left( \left(1- \frac{d}n - \frac{d}n z m(z) \right) \Sigmab - z \I_d \right)^{-1}
\]
so that for $d < n$ by taking $z = 0$ we have
\[
  m(0) = \frac{n}d \frac1{n-d} \tr \Sigmab^{-1}.
\]
The steps (ii) and (iii) follow exactly the same line of arguments as the $\bar c < 1$ case and are thus omitted.

% {\RED TODO: proof for $n>d$ sub-Gaussians. We can't appeal
% to inverse Wishart mean anymore}
% Now consider $n > d$ with $n/d \to \bar c \in (1, \infty)$.  Firstly notice
% $\beta_n = e^{d - n} \to 0$, so $\lim \mathcal{V}(\Sigmab, n) = \lim
% \frac{1}{n-d} \tr(\Sigmab^{-1})$. Noting
% $\lim \frac{1}{n-d} \tr(\Sigmab^{-1}) = \int \tau^{-1} dH(\tau)$ (where $H$ is the
% limiting spectral distribution of $\Sigmab$) is a constant, by Slutsky's theorem
% it suffices to show
% \[
%   \tr (\X^\top \X)^\dagger - \frac1{n-d} \tr \Sigmab^{-1} \to 0
% \]
% Recognizing $(\X^\top \X)^\dag = (\X^\top\X)^{-1}$ as an inverse-Wishart
% matrix when $n > d$, we have that the left-hand side is identically zero.

\subsection{Proof of \cref{c:projection}}
\label{sec:proof-of-c-projection}

Since $\X^\dagger \X = \X^\top (\X \X^\top)^\dagger \X$, to prove
\cref{c:projection}, we are interested in the limiting behavior of the
following quadratic form
\[
  \lim_{n,d \to \infty} \lim_{z \to 0^+} \frac1n \w^\top \X^\top \left( \frac1n \X \X^\top + z \I_n \right)^{-1} \X \w
\]
for deterministic $\w \in \mathbb R^{d}$ of bounded Euclidean norm (i.e., $\| \w \| \le C'$ as $n,d \to \infty$), as $n,d \to \infty$ with $n/d \to \bar c \in (0,1)$. The limiting behavior of the above quadratic form, or more generally, bilinear form of the type $\frac1n \w_1^\top \X^\top \left( \frac1n \X \X^\top + z \I_n \right)^{-1} \X \w_2$ for $\w_1, \w_2 \in \mathbb R^{d}$ of bounded Euclidean norm are widely studied in random matrix literature, see for example \cite{hachem2013bilinear}.

For the proof of \Cref{c:projection} we follow the same protocol as that of
\Cref{c:wishart}, namely: (i) we consider, for fixed $z > 0$, the limiting
behavior of $\frac1n \w^\top \X^\top \left( \frac1n \X \X^\top + z \I_n
\right)^{-1} \X \w$. Note that
\begin{align*}
  \delta(z) &\equiv \frac1n \w^\top \X^\top \left( \frac1n \X \X^\top + z \I_n \right)^{-1} \X \w = \w^\top \left( \frac1n \X^\top \X + z \I_d \right)^{-1} \frac1n \X^\top \X \w \\
  &= \| \w \|^2 - z \w^\top \left( \frac1n \X^\top \X + z \I_d \right)^{-1} \w
\end{align*}
and it remains to work on the second $z \w^\top \left( \frac1n \X^\top \X + z \I_d \right)^{-1} \w$ term.
It follows from \cite{hachem2013bilinear} that
\[
   %\delta(z) = \| \w \|^2 - z \w^\top \left( \frac1n \X^\top \X + z \I_d \right)^{-1} \w  \to \| \w \|^2 - \w^\top (\I_d + m(z) \Sigmab)^{-1} \w^\top
   z \w^\top \left( \frac1n \X^\top \X + z \I_d \right)^{-1} \w - \w^\top (\I_d + m(z) \Sigmab)^{-1} \w^\top \to 0
\]
almost surely as $n,d \to \infty$, where we recall $m(z)$ is the unique solution to \eqref{eq:def-m}.

We move on to step (ii), under the assumption that $c \le \lambda_{\min} (\Sigmab) \le \lambda_{\max} (\Sigmab) \le C$ and $\| \w \| \le C'$, we have
\begin{align*}
  \lambda_{\max} \left( \frac1n \X^\top \left( \frac1n \X \X^\top + z \I_n \right)^{-1} \X \right) &\le \frac{ \lambda_{\max} (\X \X^\top/n) }{ \lambda_{\min} (\X \X^\top/n) + z } \le \frac{ \lambda_{\max} (\Z \Z^\top/n) \lambda_{\max} (\Sigmab) }{ \lambda_{\min} (\Z \Z^\top/n) \lambda_{\min} (\Sigmab) } \\
  &\le 4 \frac{ (\sqrt{\bar c} + 1)^2 C }{ (\sqrt{\bar c} -1)^2 c }
\end{align*}
so that $\delta(z)$ remains bounded and similarly for its derivative
$\delta'(z)$, which, by Arzela-Ascoli theorem, yields uniform
convergence and we are allowed to take the $z \to 0^+$
limit. Ultimately, in step (iii) we exchange the two limits with
Moore-Osgood theorem, concluding the proof.

\subsection{Finishing the proof of Theorem~\ref{t:asymptotic}}

To finish the proof of Theorem~\ref{t:asymptotic}, it remains to write
\[
  \MSE{\X^\dagger\y} = \sigma^2\E\big[\tr\big((\X^\top\X)^{\dagger}\big)\big] +
    \w^{*\top}\E\big[\I-\X^\dagger\X\big]\w^*
\]
Since $\lambda_n = \frac{d-n}{ \tr (\Sigmab + \lambda_n \I)^{-1} }$,
by \Cref{c:wishart} and \Cref{c:projection} we have
$\MSE{\X^\dagger\y} - \Mc(\Sigmab, \w^*, \sigma^2, n) \to 0$ as $n,d
\to \infty$ with $n/d \to \bar c \in (0,\infty) \setminus \{ 1 \}$,
which concludes the proof of Theorem~\ref{t:asymptotic}.
% \begin{align*}
%   &\MSE{\X^\dagger\y} - \Mc(\Sigmab, \w^*, \sigma^2, n) \\
%   &=
%     \underbrace{\left(
%       \sigma^2\E\big[\tr\big((\X^\top\X)^{\dagger}\big)\big]
%       - \frac{\sigma^2(1 - \alpha_n)}{\lambda_n}
%     \right)}_{\to 0} + \underbrace{\left(
%       \w^{*\top}\E\big[\I-\X^\dagger\X\big]\w^*
%       - \lambda_n \w^{*\top} (\Sigmab + \lambda_n \I)^{-1} \w^*
%     \right)}_{\to 0}
% \end{align*}
%as $n,d \to \infty$ with $n/d \to \bar c \in (0,1)$.
%This allows us to conclude the proof of

% For
% the case $n <d$, note that $\X^\dagger \X = \I$ and
% \[
%   \MSE{\X^\dagger\y} - \frac1{n-d} \tr (\Sigmab^{-1}) \to 0.
% \]

\section{Additional details for empirical evaluation}
\label{a:empirical}

Our empirical investigation of the rate of asymptotic convergence in \Cref{t:asymptotic} (and, more
specifically, the variance and bias discrepancies defined in
Section~\ref{sec:asymp-conj-details}), in the context of
Gaussian random matrices, is related to open problems which have been
extensively studied in the literature. Note that when
$\X=\Z\Sigmab^{1/2}$ were $\Z$ has i.i.d.~Gaussian entries (as in
Section \ref{sec:asymp-conj-details}), then $\W=\X^\top\X$ is known as the
pseudo-Wishart distribution (also called the singular Wishart),
denoted as $\W \sim \Pc\Wc(\Sigmab, n)$, and the variance
term from the MSE can be written as $\sigma^2\E[\tr(\W^\dagger)]$.
\cite{srivastava2003} first derived the probability density function of the
pseudo-Wishart distribution, and
\cite{cook2011} computed the first and second moments of generalized
inverses. However, for the Moore-Penrose inverse and arbitrary 
covariance $\Sigmab$, \cite{cook2011} claims that the quantities required to
express the mean ``do not have tractable closed-form representation.''
The bias term, $\w^{*\top}\E[\I-\X^\dagger\X]\w^*$, has connections to
directional statistics.  Using the SVD, 
we have the equivalent representation $\X^\dagger \X = \V \V^\top$ where $\V$
is an element of the Stiefel manifold $V_{n,d}$ (i.e., orthonormal $n$-frames
in $\R^d$).  The distribution of $\V$ is known as the matrix angular central
Gaussian (MACG) distribution \citep{chikuse1990matrix}. While prior work has
considered high dimensional limit theorems \citep{CHIKUSE1991145} as well as
density estimation and hypothesis testing \citep{CHIKUSE1998188} on $V_{n,d}$,
they only analyzed the invariant measure (which corresponds in our setting to
$\Sigmab = \I$), and to our knowledge a closed form expression of
$\E[\V\V^\top]$ where $\V$ is distributed according to MACG with arbitrary
$\Sigmab$ remains an open question.

For analyzing the rate of decay of variance and bias discrepancies (as
defined in Section \ref{sec:asymp-conj-details}), it suffices to only consider diagonal
covariance matrices $\Sigmab$.  This is because if $\Sigmab = \Q \D \Q^\top$ is
its eigendecomposition and $\X\sim\Nc_{n,d}(\zero, \I_n \otimes \Q\D\Q^\top)$,
then we have for $\W \sim \Pc\Wc(\Sigmab, n)$ that $\W \overset{d}{=} \X^\top
\X$ and hence, defining $\Xt\sim\Nc_{n,d}(\zero,\I_n \otimes \D)$, by linearity
and unitary invariance of trace,
\begin{align*}
  \E[\tr(\W^\dagger)]
  &= \tr\big( \E[(\X^\top\X)^\dagger] \big)
  = \tr\Big( \Q\E\big[(\Xt^\top\Xt)^\dagger\big]\Q^\top \Big)
  = \tr\Big( \E\big[(\Xt^\top\Xt)^\dagger\big] \Big)
  = \E\left[\tr \big((\Xt^\top\Xt)^\dagger\big) \right].
\end{align*}
Similarly, we have that $\E[\X^\dagger\X]=\Q\E\big[\Xt^\dagger\Xt\big]\Q^\top$,
and a simple calculation shows that the bias discrepancy is
also independent of the choice of matrix $\Q$.

In our experiments, we increase $d$ while keeping the aspect ratio $n/d$
fixed and examining the rate of decay of the discrepancies.
We estimate $\E\big[\tr(\W^\dagger)\big]$ (for the variance) and
$\E[\I-\X^\dagger\X]$ (for the bias) through Monte Carlo sampling.
Confidence intervals are constructed using ordinary bootstrapping for
the variance. We rewrite the supremum over $\w$ in bias discrepancy as
a spectral norm: 
\[\big\|\Bc(\Sigmab,n)^{-\frac12}\E[\I-\X^\dagger\X]\Bc(\Sigmab,n)^{-\frac12} -
  \I\big\|,\]
and apply existing methods for constructing bootstrapped operator
norm confidence intervals described in \cite{lopes2019bootstrapping}.  To
ensure that estimation noise is sufficiently small, we continually increase the
number of Monte Carlo samples until the bootstrap confidence intervals are
within $\pm 12.5\%$ of the measured discrepancies.  We found that while
variance discrepancy required a relatively small number of trials (up
to one thousand), estimation noise was much larger for the bias
discrepancy, and it necessitated over two million trials to obtain
good estimates near $d=100$.

\subsection{Eigenvalue decay profiles}
\label{sec:eig-decay-details}

Letting $\lambda_i(\Sigmab)$ be the $i$th largest eigenvalue of
$\Sigmab$, we consider the
following eigenvalue profiles (visualized in Figure~\ref{fig:eig-decays}):
\begin{itemize}
  \item \texttt{diag\_linear}: linear decay, $\lambda_i(\Sigmab)= b-a i$;
  \item \texttt{diag\_exp}: exponential decay, $\lambda_i(\Sigmab) = b\,10^{- a i} $;
  \item \texttt{diag\_poly}: fixed-degree polynomial decay, $\lambda_i(\Sigmab) = (b-a i)^2$;
  \item \texttt{diag\_poly\_2}: variable-degree polynomial decay, $\lambda_i(\Sigmab) = b i^{-a}$.
\end{itemize}
The constants $a$ and $b$ are chosen to ensure $\lambda_{\text{max}}(\Sigmab) = 1$ and
$\lambda_{\text{min}}(\Sigmab) = 10^{-4}$ (i.e., the condition number
$\kappa(\Sigmab) = 10^{4}$ remains constant).