### 2022-08-23 design/pseudoinverse.tex

\documentclass[12pt]{sty/colt2019/colt2018-arxiv}
% \else
% \documentclass[anon,12pt]{sty/colt2019/colt2019} % Anonymized submission
% \documentclass[12pt]{colt2019} % Include author names
% \fi

% The following packages will be automatically loaded:
% amsmath, amssymb, natbib, graphicx, url, algorithm2e

% \title[Short Title]{Full Title of Article}
\title{Averaging subsampled pseudoinverses}
\usepackage{times}

\usepackage{color}
\usepackage{cancel}
\usepackage{wrapfig}
\usepackage{caption}
\usepackage{algorithm}
\usepackage{xfrac}
\usepackage{algorithmic}
\input{shortdefs}

\coltauthor{\Name{Micha{\l } Derezi\'{n}ski}
  \Email{mderezin@berkeley.edu}\\
  \addr Department of Statistics, UC Berkeley 
}

\begin{document}

\maketitle
For a square $n\times n$ matrix $\A$, we use $\adj(\A)$ to denote its adjugate, defined as an $n\times n$ matrix whose $(i,j)$th entry is $(-1)^{i+j}\det(\A_{-j,-i})$, where $\A_{-j,-i}$ denotes $\A$ without $j$th row and $i$th column. 
The adjugate matrix provides a key connection between the inverse and
the determinant because for any invertible matrix $\A$, we have
$\adj(\A) = \det(\A)\A^{-1}$. Also let $\pdet(\A)$ denote the pseudo-determinant.

Given a sequence of matrices $(\Z_i)_{i=1}^n$, we will use $\Z_{[S]}$ to
denote $\sum_{i\in S}\Z_i$ and let $\Z_{[n]}=\sum_{i=1}^n\Z_i$. 
Consider the following assumption, which is an
extension of the traditional \emph{general position} assumption.
\begin{assumption}
We say that a sequence of square rank-1 matrices $(\Z_i)_{i=1}^n$ is in
$r$-general position if for any subset $S\subseteq[n]$ of size at
least $r$ we have $\rank(\Z_{[S]}) = r$.
\end{assumption}

The fundamental result we use does not require the above assumption.
\begin{theorem}\label{t:adj}
  If $(\Z_i)_{i=1}^n$ are $d\times d$ matrices then for any $k\geq
  d-1$,
  \begin{align*}
    \sum_{S:|S|=k}\adj(\Z_{[S]})=   {n-d+1\choose k-d+1}\adj(\Z_{[n]}).
  \end{align*}
\end{theorem}

The following Theorem is a straightforward extension of the result
from \cite{unbiased-estimates-journal}, and follows as a corollary of
Theorem \ref{t:adj}.
   \begin{theorem}\label{t:sqinv}
Suppose that $(\Z_i)_{i=1}^n$ are $d\times d$ and in $d$-general
position. Then for any $k\geq d$, 
     \begin{align*}
\sum_{S:|S|=k} w_S \,\Z_{[S]}^{-1} =
       \tfrac{n-d+1}{k-d+1}\,\Z_{[n]}^{-1}\quad\text{and}\quad 
     \sum_{S:|S|=k}\!w_S = 1,\quad \text{for}\quad w_S=
\tfrac{\det(\Z_{[S]})}{{n-d\choose k-d}\det(\Z_{[n]})}.
     \end{align*}
   \end{theorem}
Note that we do not require that any matrices be psd, but we do
require invertibility. If the matrices are not psd, then some of the
weights $w_S$ will be negative, but they will still sum to 1.
Next, we relax the invertiblity assumption.

   \begin{theorem}\label{t:sqinv}
Suppose that $(\Z_i)_{i=1}^n$ are square and in $r$-general
position. Then for any $k\geq r$, 
     \begin{align*}
\sum_{S:|S|=k} w_S \,\tr(\Z_{[S]}^{\dagger}) =
       \tfrac{n-r+1}{k-r+1}\,\tr(\Z_{[n]}^{\dagger})\quad\text{and}\quad 
     \sum_{S:|S|=k}\!w_S = 1,\quad \text{for}\quad w_S=
\tfrac{\pdet(\Z_{[S]})}{{n-r\choose k-r}\pdet(\Z_{[n]})}.
     \end{align*}
   \end{theorem}
Note that trace could be replaced by any linear function of the eigenvalues
of the matrix. 
     \begin{proof}
Let $\X,\Y$ be $n\times d$ matrices such that
$\Z_{[S]}=\X_S^\top\Y_S$, where $\X_S$ selects out the rows of
$\X$. Note that $\L=\X\Y^\top$ has rank $r$, so there are $n\times r$
matrices $\A,\B$ such that $\Y\X^\top=\B\A^\top$. Moreover, since
       $\Y_S\X_S^\top=\L_{S,S}=\B_S\A_S^\top$ for any $S\subseteq [n]$, we have
       \begin{align*}
         \tr\big(\Z_{[S]}^\dagger\big) =
         \tr\big((\Y_S\X_S^\top)^\dagger\big) =
         \tr\big((\B_S\A_S^\top)^\dagger\big) = \tr\big((\A_S^\top\B_S)^{-1}\big).
       \end{align*}
       In particular, we also have $\tr\big((\X^\top\Y)^\dagger\big) =
       \tr\big((\A^\top\B)^{-1}\big)$. The result now follows by
       applying Theorem \ref{t:sqinv} to matrices $\tilde{\Z}_{[S]}=\A_S^\top\B_S$.
     \end{proof}

%\bibliographystyle{alpha}
\bibliography{pap}

\end{document}