### 2022-08-23 design/talk/Michal_Derezinski_Answers.tex

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{fullpage}
% \usepackage[1in]{geometry}                		% See
%geometry.pdf to learn the layout options. There are lots.

%\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{url}
%SetFonts

%SetFonts

\title{Improved guarantees and a multiple-descent curve for \\
  % the CSSP
  Column Subset Selection %Problem
  and the Nystr\"om method}
\author{Micha{\l} Derezi\'nski (speaker), Rajiv Khanna and Michael W. Mahoney}
 \date{}							% Activate to display a given date or no date

\begin{document}
\maketitle

\end{document}