### 2022-08-23 design/talk/asymptotic.tex

\documentclass[10pt]{beamer}
%\beamertemplateshadingbackground{brown!70}{yellow!10}
\mode<presentation>
{
  %\usetheme{Warsaw}
  \usecolortheme{crane}
  % or ...

%  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]{}
\usepackage{tikz,pgfplots}
\pgfplotsset{compat=newest}
\usepackage[utf8]{inputenc}
\usetikzlibrary{patterns}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{colortbl}
%\usepackage{multicol}
\usepackage{cancel}
\usepackage{ulem}
\usepackage{multirow}
\usepackage{relsize}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{forloop}% http://ctan.org/pkg/forloop
\newcounter{loopcntr}
\newcommand{\rpt}[2][1]{%
  \forloop{loopcntr}{0}{\value{loopcntr}<#1}{#2}%
}
%\pagestyle{plain}
%\input{defs2}
\input{../shortdefs}

\edef\polishl{\l}
\setlength{\columnsep}{0.7em}
\setlength{\columnseprule}{0mm}
\setlength{\arrayrulewidth}{1pt}

\newcommand{\svr}[1]{{\textcolor{darkSilver}{#1}}}
\definecolor{brightyellow}{cmyk}{0,0,0.7,0.0}
\definecolor{lightyellow}{cmyk}{0,0,0.3,0.0}
\definecolor{lighteryellow}{cmyk}{0,0,0.1,0.0}
\definecolor{lightestyellow}{cmyk}{0,0,0.05,0.0}
\AtBeginSection[]
{
\begin{frame}<beamer>
\frametitle{Outline}
\tableofcontents[currentsection]
\end{frame}
}
\def\layersep{2.5cm}

%  \fboxsep=3pt
% %\fboxsep=0mm%padding thickness
% \fboxrule=2pt%border thickness

\setkeys{Gin}{width=0.7\textwidth}

\title[]{Strong asymptotic equivalence}

\author[]{Micha{\l} Derezi\'{n}ski\\
UC Berkeley}

\begin{document}
\begin{frame}
  \titlepage
\end{frame}

\linespread{1.3}

\begin{frame}
  \frametitle{Standard asymptotic equivalence}
  Matrix sequences $\A_n$, $\B_n$ (deterministic or random) of growing dimensions are
  \emph{equivalent}, denoted by
  \begin{align*}
    \A_n\asymp \B_n,
  \end{align*}
  if
  \begin{align*}
    \lim_{n\rightarrow\infty}\big|\tr\,\C_n(\A_n-\B_n)\big| =0,
  \end{align*}
  almost surely, for any sequence $\C_n$ of not necessarily symmetric matrices with bounded trace
  norm, i.e., such that
  \begin{align*}
    \lim\sup\|\C_n\|_{\tr}<\infty.
  \end{align*}
  Here, the trace norm is defined by $\|\M\|_{\tr} = \tr((\M^\top\M)^{\frac12})$.
\end{frame}

\begin{frame}
  \frametitle{\emph{New:} Strong asymptotic equivalence}
  Positive semi-definite matrix sequences $\A_n$, $\B_n$ (deterministic or random) of growing dimensions are
  \emph{strongly equivalent}, denoted by
  \begin{align*}
    \A_n\simeq \B_n,
  \end{align*}
  if
  \begin{align*}
    \A_n^{\dagger/2}\B_n\A_n^{\dagger/2} \asymp
    \B_n^{\dagger/2}\A_n\B_n^{\dagger/2}\asymp \I_n,
  \end{align*}
  where $\I_n$ is the sequence of identity matrices with matching
  dimensions, and $\M^{\dagger/2}=(\M^\dagger)^{\frac12}$ is the
  square-root of the Moore-Penrose inverse.\\[5mm]
  \emph{Alternative condition:} \\
  Instead of relying on standard equivalence, we can use the spectral norm:
  \begin{align*}
    \lim_{n\rightarrow\infty}\|\A_n^{\dagger/2}\B_n\A_n^{\dagger/2} -
    \I_n\|
    =
        \lim_{n\rightarrow\infty}\|
    \B_n^{\dagger/2}\A_n\B_n^{\dagger/2} - \I_n\|
    = 0.
  \end{align*}
\end{frame}

\begin{frame}
  \frametitle{Desired properties of strong equivalence}
  \begin{enumerate}
  \item Strong equivalence implies standard equivalence, i.e.:
    \begin{align*}
      \text{if}\quad \A_n\simeq\B_n \quad\text{then}\quad
      \A_n\asymp\B_n
      \end{align*}
  \item if $\A_n$ and $\B_n$ have uniformly bounded condition numbers,
    then equivalence implies strong equivalence\\[3mm]
  \item Some form of calculus of deterministic equivalents applies to
    the strong equivalence relation.
  \end{enumerate}
  \vspace{3mm}

  \emph{Note:} Strong equivalence differs from standard equivalence
  when sequences $\A_n$ and $\B_n$ have vanishing eigenvalues.
\end{frame}

\begin{frame}
  \frametitle{Some examples}

Let $\X_n\sim\Nc(\zero,\I_n\otimes\Sigmab_n)$. We can consider various sequences:
\begin{align*}
  \A_n &= \E[\X_n^\dagger\X_n],\\
  \A_n &= \E[(\tfrac1n\X_n^\top\X_n + \lambda_n\I)^{-1}],\quad\text{where}\quad
  \lambda_n\rightarrow 0,\\
  \A_n &= \E[(\tfrac1n\X_n^\top\X_n + \lambda_n\I)^{-p}],\quad\text{where}\quad
  \lambda_n\rightarrow 0
\end{align*}
\emph{Note:} Standard deterministic equivalents for $\A_n$ are $\B_n=(\gamma_n\Sigmab + \delta_n\I)^{-p}$, for the right choices
of $\gamma_n,\delta_n$\\[3mm]

However, they will not satisfy the \emph{strong} equivalence when, e.g.:
\begin{align*}
  \Sigmab_n =
  \begin{pmatrix}
    \I_n&\zero\\
    \zero & \frac1n\I_n
  \end{pmatrix}.
\end{align*}
\emph{Question:} How do standard and strong equivalences differ when
$\Sigmab_n$ exhibits some form of polynomial or exponential spectral decay?

\end{frame}

\end{document}