### 2022-08-23 design/talk/talk-nystrom-script.tex

\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=.5in]{geometry}
\usepackage{xcolor}
\usepackage{url}
%\usepackage[sc]{mathpazo}
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
\newcommand{\Hrule}{\rule{\linewidth}{0.3mm}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\makeatletter% since there's an at-sign (@) in the command name
\renewcommand{\@maketitle}{%
  \parindent=0pt% don't indent paragraphs in the title block
  \centering
  {\Large \bfseries\textsc{\@title}}
  \HRule\par%
  \textbf{\@author \hfill \@date}
  \par
}
\def\b{{~\textbullet~}}
\makeatother% resets the meaning of the at-sign (@)

\title{Nystr\"om Talk Script}
\author{Micha{\l} Derezi\'{n}ski}
\date{\texttt{mderezin@berkeley.edu}}

\begin{document}
  \maketitle% prints the title block
\setlength{\parskip}{0.5\baselineskip}
\setlength{\parindent}{0pt}

Hi, my name is Micha{\l}, and I will be talking about an intriguing
phenomenon, which we refer to as the multiple-descent curve in subset
selection. This is joint work with Rajiv and Michael.

A core motivation for the tasks we are studying is interpretable
dimensionality reduction. As an example, consider a biological
application, with a large genetics dataset where each individual is
described by a sequence of genetic variations, called Single
Nucleotide Polymorphisms, SNPs for short. Prior work has shown that
such data can be effectively characterized by a subset of SNPs, chosen
so that it is correlated with the top Principal Components of the data
matrix. This subset of SNPs can then be used by domain experts to
extract insight from the data. Many other applications of
interpretable dimensionality reduction have been demonstrated in the
literature, including for example document clustering.

The two basic dimensionality reduction tasks we consider in this work
are Column Subset Selection and the
Nystrom method. Those two tasks are very closely related, even though they often
appear independently in the literature. The first one, on the left, is
most commonly described as feature selection from a data matrix A,
although we could just as well consider a setting where the columns of
A are the instances instead of features. For the Nystrom method, on
the right, the matrix in question is a square kernel matrix, and we
wish to represent it by a subset of entries. This task arises in the
context of kernel machines, Gaussian Process regression and
Independent Component Analysis, among others. As we demonstrate in the
paper, an equivalence between these tasks can be established, which
means that all of our results apply in both settings.

In low-rank approximation our goal is to project data onto a
low-dimensional subspace, while minimizing the projection
error. Without additional constraints, there is a unique choice to
minimize this error, by projecting onto the top k principal components
in the data. However, typically these principal component vectors will
not correspond to any interpretable feature or instance, which is why
we may want to\b choose a subspace spanned by a column subset of the
data matrix. Naturally, there are other notions of interpretability
than selecting subsets of the data matrix, however this model provides
a useful framework for analyzing, what can be described as, the cost
of interpretability. We measure this cost via the so-called
approximation factor, comparing the error of the best interpretable
representation, relative to the optimal non-interpretable choice given
by the principal components, for a given dimensionality parameter k.

A large body of literature has been dedicated to obtaining worst-case
bounds on the approximation factor that describes the cost of
interpretability. In particular, it was shown, for 
Column Subset Selection by Deshpande, Rademacher, Vempala and Wang,
and for the Nystrom method by Belabbas and Wolfe, that the cost of
interpretability in rank k approximation can always be bounded by a
factor of k+1. These results are still used as a technical tool
in recent works, including the best paper at the 2019 International
Conference on Machine Learning, which
applied them to sparse variational Gaussian Process regression. The
above bound is in fact worst-case optimal in that there are problem
instances for which factor k+1 cannot be improved. Nevertheless, these
results are still somewhat unsatisfying because they suggest that the
cost of interpretability gets progressively higher as the rank
parameter k grows, as seen in the plot below, which is not generally
observed in practice. To understand this,\b we empirically tested the
approximation factor for the worst-case example given by Deshpande et
al. As we can see, the factor does reach the worst-case bound for one
value of k, but then it rapidly drops down in either direction. So,
even for the worst-case example, a high approximation factor is only a
corner case. To explain this,\b we go beyond worst-case analysis, and
our bounds are able to accurately capture the non-linear behavior of
the approximation factor.

In particular, our new guarantees show that the worst-case example is
part of a larger phenomenon, which we refer to as the multiple-descent
curve in subset selection. This curve is illustrated in the
following plot, both empirically and through our theory. As we can
see, the approximation factor exhibits occasional sharp spikes, but is
otherwise very small.  Our new lower bound shows that more than one
spike can occur for one data matrix, which was not the case for the
previously demonstrated worst-case constructions. Moreover, we show
that the spikes are correlated with sharp drops in the data spectrum.

The multiple-descent curve in subset selection is very much
reminiscent of the so-called double descent curve, proposed by Belkin
et al, which is exhibited by the generalization error of many machine
learning models, when varying the ratio between the number of
parameters and the training data size. In this context, we can
distinguish two regimes: the Classical one, where we have more data
than parameters, and the Modern one, which includes deep learning
architectures, where there is more parameters than data and so the
model can interpolate the training set. The phase transition between
the two regimes exhibits a peak in the generalization
error. Naturally, there are important differences compared to our
setting: first, we consider a combinatorial optimization task of
subset selection, as opposed to a statistical learning one, and
second, we study the approximation factor instead of the
generalization error. However, several connections arise.\b Similarly as
in the case of multiple-descent in subset selection, the double
descent peak is correlated with the condition number of the data, and
moreover, recent works showed that the generalization error can also
exhibit multiple peaks. Finally, techniques similar to ours can be
used to obtain exact expressions for double descent in terms of the
mean squared error of the minimum norm interpolating linear model,
which is actually what we used to generate the plot shown on the
right.

So, what properties of the data affect the non-linear behavior of the
cost of interpretability? For example, 
do we expect the cost to be smaller if the data
is more or less uniformly distributed in the domain space, as in
Dataset 1, or if it has some low-dimensional structure, as in Dataset 2?
Well, first, we need to be able to easily characterize such properties of the data.\b For
this we can use the so-called ``spectral decay profile'', which looks at
the magnitude of the largest directions of variance in the data, represented
mathematically as the singular
values of the data matrix. To illustrate the structure of the dataset through spectral decay, we plot
the top singular values in decreasing order. For example, for the first dataset, we see a relatively
flat decay with multiple singular values of comparable
magnitude, because there are many directions with large
variance, whereas for the second dataset, there is one or two large singular values
followed by a rapid decay. This corresponds to the
low-dimensional structure we can see in the top right.

There are several types of spectral decays which are commonly
observed in real world datasets, such as a flat decay, a sharp
drop, and a  range of smooth decays with polynomial or
exponential rates. In fact, often we may be able to anticipate
the type of spectral decay even before we look at the data, for
example because of some prior knowledge coming from domain
experts about the underlying data distribution, and in some cases
we can actually control it through model selection, by choosing
an apprioriate similarity kernel.

In this work, we used the spectral decay profile of the data to
provide a complete characterization of the cost of interpretability,
showing that if the data exhibits sufficiently 
smooth spectral decay, then the spikes in the multiple-descent curve
can be avoided. In particular, this  
allows us to deduce that the spikes in the cost of
interpretability are directly correlated with sharp drops in the
spectrum of the data. This also implies that when the spectral decay is
sufficiently smooth, for example, when it exhibits a polynomial
rate which is commonly observed in practice, then we can prove
that there is no spikes and the cost of interpretability is
small for all values of the dimensionality parameter,
significantly improving on the prior work.

Of course, the practical performance of dimensionality reduction depends on
the subset selection method used. Here, we focus primarily on the method called
Determinantal Point Process sampling. % , although in the paper we
% empirically demonstrate that the multiple-descent curve is also
% exhibited by another commonly used method, called greedy subset
% selection.
A Determinantal Point Process is a non-i.i.d. randomized
method that uses negative correlation to induce diversity in the
sampled subset. DPPs have found many applications in machine
learning for the past 15 years or so, including in recommendation
systems, experimental design and stochastic optimization, and they can
be implemented very efficiently, with runtimes faster than singular
value decomposition. For a state-of-the-art implementation, check out
a paper by myself, Daniele Calandriello and Michal
Valko from the last year's Neurips conference. And to learn more about
related applications of DPPs, see a new 
survey by myself and Michael Mahoney, as well as an earlier monograph
by Kulesza and Taskar.

In our empirical results, we showed that the multiple-descent
phenomenon can be observed not only in worst-case 
examples but also in real-world subset selection tasks. Here, we
demonstrate this for the Nystrom method with the Gaussian RBF kernel,
studying the effect of varying the RBF parameter. On the top plots, we
show the Nystrom approximation factor for two Libsvm datasets whereas
at the bottom we compare it to decreasing singular values of the
data. We can clearly see that the spikes in approximation factor align
with big drops in the spectral decay, and that this behavior is
greatly influenced by the value of the RBF parameter. Decreasing this
parameter results in a smoother spectral decay, which in turn
eliminates the spikes, resulting in a uniformly small approximation
factor. The right choice of the RBF parameter ultimately depends on
the down-stream machine learning task, so depending on the structure
of the problem, we may end up in the regime where the
approximation factor exhibits a multiple-descent curve or in the
regime where it is relatively flat. Here, I showed the results when using a
Determinantal Point Process for subset selection, however, remarkably,
we can observe the same phenomenon even if we switch to a completely
different method, called Greedy Subset Selection.\b This suggests that
the multiple-descent curve is not specific to DPPs, but rather
it is a method-agnostic phenomenon.

So, to conclude, in this work we studied the cost of interpretability
in dimensionality reduction. We particularly focused on two methods of
dimensionality reduction, namely, Column Subset Selection and the
Nystrom method, both of which use data subsets to construct small
interpretable representations of large datasets. In our main results,
we overcome the limitations of prior worst-case analysis to accurately
capture the non-linear behavior of the cost of interpretability. Moreover, our
analysis reveals that the previously observed worst-case examples are
part of a larger phenomenon, which we call the multiple-descent curve
in subset selection. This multiple-descent curve can be observed not
only in artificially constructed pathological examples but also in
real-world problems. However, it can generally be avoided by tuning
model parameters.

Here is a complete list of references from the talk, and thank you
very much for your attention!

\newpage

\end{document}