### 2022-08-23 ic/aistats2021/aistats2021_author_response.tex

\documentclass{article}

\usepackage{aistats2021_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % define colors in text
\usepackage{xspace}         % fix spacing around commands
\usepackage{mathtools,natbib}

\setlength{\bibsep}{0.0pt}

\bibliographystyle{abbrv}

\begin{document}

We thank our reviewers and will fix all typos/grammar errors. Any
reproducability issues will be addressed when we release all implementation
(reviewer 1) and experiment code in early 2021. We will also add additional
details on $n$-schools notation and motivation (reviewers 2, 6).

\textbf{Reviewer 2}:

We respectfully point out that reviewer 2 may have entirely misunderstood our work as
posterior adaptation rather than amortized inference and we strongly recommend
reconsideration of the claims in this review.
While reviewer 2 summarizes the ``first step'' of LIC as running
MCMC to sample $p(x \mid y)$, in fact LIC (see equation 2) and other methods
for amortized inference \cite{le2017inference} sample the joint distribution
$p(x,y)$ by running the probabilistic program forwards. This misunderstanding
is further evidenced by reviewer 2's major concern on lack of novelty
compared against three prior works on posterior adaptation
\cite{de2001variational,neklyudov2018metropolis,titsias2019gradient}, when in
fact our work is on amortized inference (see section 1.2). As amortized
inference must provide speedups (i.e. amortize the upfront compilation
cost) across many different observations $y$ and not just a single target
posterior $p(x \mid y)$, the prior works in reviewer 2's first major concern
are not comparable to our work. Prior work in IC \cite{le2017inference} also
does not make direct comparison against posterior adaptation methods, and we have
followed suit.
Additionally, reviewer 2's second major concern related to being ``safe from
mode-collapse'' are likely a consequence of misunderstanding the ``first
step'' as performing MCMC sampling of the posterior. Because no Markov chains
are actually involved during learning for LIC, there is no concern of any chains getting stuck
/ collapsing.

\textbf{Reviewer 3}:

The major criticism from the reviewer in ``Appropriateness of Graph Neural Networks in theory''
is that averaging together nodes in a Markov blanket (e.g. the mean and variance) may
not yield meaningful signal. While this is true for a 1D-embedding, we point out that
the vector space embeddings (``embedding net''s in Figure 2) yield multi-dimensional
embeddings where different dimensions handle how nodes in the Markov blanket ``play
different roles from one another.'' For the mean/variance example, the
embedding networks could potentially learn to embed the mean to dimension 1
and variance to dimension 2 so the resulting sum yields a 2-dimensional
[mean, variance] vector.

Regarding weakness of empirical validation, we attempted to conduct
the first experiment of the two ``missing'' experiments
and found LIC to improve over IC \cite{le2017inference} in terms of 
compilation/inference times, model size, and ESS/PLL on Bayesian logistic
regression and $n$-schools. 
However, we opted to not include these results because IC's reference
implementation (\texttt{pyprob}, \cite{le2017inference}) lacks support for
vector-valued random variables (see e.g.
\texttt{pyprob.distributions.mixture.Mixture.event\_shape}) so we were forced
to unvectorize our \texttt{pyprob} models. If there is interest, we can
include these results in the appendix.

Regarding relation to prior work, we concede that handling the open-universe
case is not novel and will remove this claim from our paper. However,
although neural Gibbs proposers are also used in \cite{wang2018meta}, a key
distinction is that they ``focus on hand-identified common structures'' to
construct proposers whereas LIC (and related work in IC) automate proposal
construction. While their framework includes Markov blanket parameterized
proposers, they raise the issue that ``Markov blankets... might not be
consistent across all instantiations'' whereas we use permutation-invariance
to motivate graph neural networks as a solution for summarizing dynamic
Markov blankets. We will include this discussion in related works.

With respect to additional questions:
    Infinite-domain discrete random variables -- yes, and we are
    actively looking at $q(\cdot; \phi) \sim \text{Poisson}(\phi)$ and
    negative-binomial.
    ``Is $x_i$ actually part of the input?'' -- yes, see Figure 2.
    Imperative vs declarative -- agree, will fix in camera-ready.
    On reproduceability -- GMM components is on page 7.
\bibliography{response}

\end{document}