### 2022-11-07 intro.tex
Updates from Overleaf

\chapter{Introduction}

In \todo{Chapter design}, we consider approximately optimal Bayesian experimental design
using an adaptive row sampling algorithm based on DPPs and show that it provides
good approximations. Through generalizing the proof techniques \todo{Determinant preserving random matrices}
applied here, in \todo{Chapter double descent} an extension of the DPP-based design is analyzed in
closed-form for the over-parameterized $n < d$ regime and predicts a double-descent phenomenon in
linear regression which closely matches empirical experiments.
In \todo{Chapter random projections} we isolate the part of the proof involving concentration
of bilinear forms of matrix resolvents away from the DPP-based design in order to obtain bounds on
expected projections $X_S^\dag X_S$ when $X_S = S X$ is obtained by sub-Gaussian sketching matrix $S$.

The second three chapters concern the setting of probabilistic programming, a computational tool
for specifying probabilistic models and automating inference through MCMC and VI. In particular, we  
study approximating a target density $p(x)$ with approximations $q(x)$ for the purposes of
importance sampling and variational inference.
In \todo{Chapter LIC}, we consider parameterizing $q(x)$ by graph neural networks which condition each
node on its Markov blanket. This reduces the conditioning sets for a node, resulting in improvements over
\cite{le2017inference} and run-times which depend on sparsity in the graphical model rather than
the length of execution traces.
When the target $p(x)$ is both multivariate and heavy tailed, \todo{Chapter FTVI}
considers the problem of tail anisotropy through both a theoretical and practical perspective.
We establish that prior fat-tailed estimators \citep{jaini2020tails} are tail isotropic, propose
an anisotropic approximation (fat-tailed variational inference, FTVI) where
an anisotropic product base measures is pushed forwards through a bijective neural network,
and confirm that in practice FTVI improves both density estimation as well as variational inference.
However, we find in practice FTVI and other tail-adaptive approximations often have trouble optimizing
the tail parameter. In \todo{Chapter HTA}, we consider addressing this issue during static analysis of a
probabilistic program's source code. We define generalized Gamma tail asymptotics for a number of
elementary distributions and establish how the tail asymptotics are transformed under algebraic transformations
such as sums and products. This enables a priori computation of tail parameters, which we show
improves the stability and convergence of a number of inference tasks.