### 2022-11-07 abstract.tex
Adds devcontainer and magic comments

% !TeX root = thesis.tex
% (This file is included by thesis.tex; you do not latex it by itself.)

\begin{abstract}

    % The text of the abstract goes here.  If you need to use a \section
    % command you will need to use \section*, \subsection*, etc. so that
    % you don't get any numbering.  You probably won't be using any of
    % these commands in the abstract anyway.

    In \cref{ch:design}, we consider approximately optimal Bayesian experimental design
    using an adaptive row sampling algorithm based on DPPs and show that it provides
    good approximations. Through generalizing the proof techniques \todo{Determinant preserving random matrices}
    applied here, in \cref{ch:double_descent} an extension of the DPP-based design is analyzed in
    closed-form for the over-parameterized $n < d$ regime and predicts a double-descent phenomenon in
    linear regression which closely matches empirical experiments.
    In \cref{ch:projections} we isolate the part of the proof involving concentration
    of bilinear forms of matrix resolvents away from the DPP-based design in order to obtain bounds on
    expected projections $X_S^\dag X_S$ when $X_S = S X$ is obtained by sub-Gaussian sketching matrix $S$.

    The second three chapters concern the setting of probabilistic programming, a computational tool
    for specifying probabilistic models and automating inference through MCMC and VI. In particular, we
    study approximating a target density $p(x)$ with approximations $q(x)$ for the purposes of
    importance sampling and variational inference.
    In \cref{ch:lic}, we consider parameterizing $q(x)$ by graph neural networks which condition each
    node on its Markov blanket. This reduces the conditioning sets for a node, resulting in improvements over
    \cite{le2017inference} and run-times which depend on sparsity in the graphical model rather than
    the length of execution traces.
    When the target $p(x)$ is both multivariate and heavy tailed, \cref{ch:ftvi}
    considers the problem of tail anisotropy through both a theoretical and practical perspective.
    We establish that prior fat-tailed estimators \citep{jaini2020tails} are tail isotropic, propose
    an anisotropic approximation (fat-tailed variational inference, FTVI) where
    an anisotropic product base measures is pushed forwards through a bijective neural network,
    and confirm that in practice FTVI improves both density estimation as well as variational inference.
    However, we find in practice FTVI and other tail-adaptive approximations often have trouble optimizing
    the tail parameter. In \cref{ch:hta}, we consider addressing this issue during static analysis of a
    probabilistic program's source code. We define generalized Gamma tail asymptotics for a number of
    elementary distributions and establish how the tail asymptotics are transformed under algebraic transformations
    such as sums and products. This enables a priori computation of tail parameters, which we show
    improves the stability and convergence of a number of inference tasks.

\end{abstract}