### 2022-11-23 intro.tex

\documentclass[thesis.tex]{subfiles}

\begin{document}

\chapter{Introduction}
\label{ch:intro}

When faced with difficult problems, randomized approximations are a tool
commonly employed by quantitative scientists which can offer algernative
algorithms and analyses. In this dissertation, we consider a range of problems
at the intersection of randomized methods and statistics.
Throughout our work, randomized methods are a unifying theme used in the first section to
construct tractable approximations and later as tools for automating computational Bayesian statistics.

Initially in the first three chapters, we develop novel analyses to recent problems
in experimental design, double descent, and random projections by using determinantal
point processes (DPP).
In \cref{ch:design}, we consider approximately optimal Bayesian experimental design
using an adaptive row sampling algorithm based on DPPs and show that it provides
good approximations. Through generalizing the previous chapter's proof techniques,
in \cref{ch:double_descent} an extension of the DPP-based design is analyzed in
closed-form for the over-parameterized $n < d$ regime and predicts a double-descent phenomenon in
linear regression which closely matches empirical experiments.
In \cref{ch:projections} we isolate the part of the proof involving concentration
of bilinear forms of matrix resolvents away from the DPP-based design in order to obtain bounds on
expected projections $X_S^\dag X_S$ when $X_S = S X$ is obtained by sub-Gaussian sketching matrix $S$.
These chapters correspond to the following publications:
\begin{itemize}
    \item \fullcite{bayesian-experimental-design}
    \item \fullcite{surrogate-design}
    \item \fullcite{precise-expressions}
\end{itemize}

In the later chapters we focus on probabilistic programming and developing theory and tools
to automate statistical inference using randomized algorithms based on
Monte Carlo Markoc chain (MCMC) and variational inference (VI).
The second three chapters concern the setting of probabilistic programming, a computational tool
for specifying probabilistic models and automating inference through MCMC and VI. In particular, we
study approximating a target density $p(x)$ with approximations $q(x)$ for the purposes of
importance sampling and variational inference.
In \cref{ch:lic}, we consider parameterizing $q(x)$ by graph neural networks which condition each
node on its Markov blanket. This reduces the conditioning sets for a node, resulting in improvements over
\cite{le2017inference} and run-times which depend on sparsity in the graphical model rather than
the length of execution traces.
When the target $p(x)$ is both multivariate and heavy tailed, \cref{ch:ftvi}
considers the problem of tail anisotropy through both a theoretical and practical perspective.
We establish that prior fat-tailed estimators \citep{jaini2020tails} are tail isotropic, propose
an anisotropic approximation (fat-tailed variational inference, FTVI) where
an anisotropic product base measures is pushed forwards through a bijective neural network,
and confirm that in practice FTVI improves both density estimation as well as variational inference.
However, we find in practice FTVI and other tail-adaptive approximations often have trouble optimizing
the tail parameter. In \cref{ch:hta}, we consider addressing this issue during static analysis of a
probabilistic program's source code. We define generalized Gamma tail asymptotics for a number of
elementary distributions and establish how the tail asymptotics are transformed under algebraic transformations
such as sums and products. This enables a priori computation of tail parameters, which we show
improves the stability and convergence of a number of inference tasks.
These chapters correspond to the following publications:
\begin{itemize}
    \item \fullcite{lic}
    \item \fullcite{ftvi}
    \item \fullcite{gga}
\end{itemize}

\end{document}